### 2019-09-08 analysis/001_kb_rp_pretest.Rmd
initial upload

---
title: "Reprint Authors per Country"
output: github_document
bibliography: pubs.bib
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE
)
```

```{r setup, echo = FALSE}
require(tidyverse)
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar")
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    Sys.getenv("kb_user"),
    Sys.getenv("kb_pwd")
  ) 
```

## Motivation

Country information about corresponding authors play a crucial role in open access funding [@Schimmer_2015]. The Web of Science contains data about this author role including affiliations, making it a valuable source to determine the productivity of corresponding authors per country ad journal.

The aim of this work is to better understand how countries of affiliation from corresponding authors are represented in the WoS-KB, and how they can be retrieved from it. Such an analysis will yield valuable insights about how to calculate the number of scholarly articles per country and journal from the WoS-KB database using the corresponding author addresses only.

## Potential issues

1. A corresponding author lists at least two different countries
2. More than one corresponding author is listed
3. Corresponding authors work in different countries

## Example 1

A corresponding author lists at least two different countries, e.g., `UT=000400754000138`

```{sql connection=jdbcConnection}
select  distinct wos_b_2019.institutions.countrycode,
        wos_b_2019.items_authors_institutions.type,
        wos_b_2019.items_authors_institutions.fk_authors,
        wos_b_2019.items.ut_eid,
        wos_b_2019.items.pubyear
from wos_b_2019.items
inner join
        wos_b_2019.items_authors_institutions                                 
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions  
where wos_b_2019.items.ut_eid = '000400754000138'
order by wos_b_2019.items_authors_institutions.fk_authors
```

Potential issues: Only one affiliation is tagged as `RP` in `wos_b_2019.items_author_institutions`.

## Example 2

More than one corresponding author is listed. For `UT_EID=000372645900002`, the Web of Science lists seven corresponding authors and four reprint adresses.

```{sql connection=jdbcConnection, max.print = NA}
select  distinct wos_b_2019.institutions.countrycode,
        wos_b_2019.items_authors_institutions.type,
        wos_b_2019.items_authors_institutions.fk_authors,
        wos_b_2019.items.ut_eid,
        wos_b_2019.items.pubyear
from wos_b_2019.items
inner join
        wos_b_2019.items_authors_institutions                                 
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions  
where wos_b_2019.items.ut_eid = '000372645900002'
order by wos_b_2019.items_authors_institutions.fk_authors
```

Potential issue: As in Example 1, Only one affiliation is tagged as `RP` in `wos_b_2019.items_author_institutions`.

## Example 3

Corresponding authors work in different countries. For `UT_EID=000372645900002`, the Web of Science lists two corresponding authors, one affilated with a Spanish, the other with a Portuguese institution.

```{sql connection=jdbcConnection}
select  distinct wos_b_2019.institutions.countrycode,
        wos_b_2019.items_authors_institutions.type,
        wos_b_2019.items_authors_institutions.fk_authors,
        wos_b_2019.items.ut_eid,
        wos_b_2019.items.pubyear
from wos_b_2019.items
inner join
        wos_b_2019.items_authors_institutions                                 
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions  
where wos_b_2019.items.ut_eid = '000389110200022'
order by wos_b_2019.items_authors_institutions.fk_authors
```

Potential issues: As in Example 1 and 2, only one affiliation is tagged with RP in `wos_b_2019.items_author_institutions`. Moreover, checking the WoS raw data reveals that the RP-author `14734495` is not affiliated with an Portuguese institution at all.

## SQL strategy to circumvent these issues

### Steps

1. Get all `FK_AUTHORS` tagged as `RP`, and store the result `rp_author_table`
2. From authors listed in `rp_author_table` obtain all countries of affilation and store them in `rp_countries_table`
3. Aggregate `rp_countries_table`

#### 1. Get all `FK_AUTHORS` tagged as `RP`, and store the result `rp_author_table`

```{sql connection=jdbcConnection, output.var="rp_authors"}
select
        distinct wos_b_2019.items_authors_institutions.fk_items,
        wos_b_2019.items_authors_institutions.fk_authors      
    from
        wos_b_2019.items      
    inner join
        wos_b_2019.databasecollection              
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items        
    inner join
        wos_b_2019.items_authors_institutions                                   
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
    inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions      
    where
        wos_b_2019.databasecollection.edition_value in ('WOS.SCI', 'WOS.SSCI', 'WOS.AHCI')                  
        and wos_b_2019.items.doctype in ('Article', 'Review')                   
        and wos_b_2019.items.pubyear in (2014, 2015, 2016, 2017)         
        and wos_b_2019.items.ut_eid in ('000389110200022', '000372645900002', '000400754000138') 
        and wos_b_2019.items_authors_institutions.type = 'RP'
```

Store in personal table space

```{r}
dbWriteTable(conn = jdbcConnection, 
             name = "rp_author_table", 
             value = rp_authors)
```

#### From authors listed in `rp_author_table` obtain all countries of affilation and store them in `rp_countries_table`

```{sql connection=jdbcConnection, output.var="rp_countries"}
select
        countrycode,
        wos_b_2019.items_authors_institutions.type,
        wos_b_2019.items_authors_institutions.fk_institutions,
        wos_b_2019.items_authors_institutions.fk_authors,
        wos_b_2019.issues.issn,
        ut_eid,
        wos_b_2019.items.pubyear
    from
        wos_b_2019.items      
    inner join
        wos_b_2019.issues                           
            on wos_b_2019.issues.pk_issues = wos_b_2019.items.fk_issues         
    inner join
        wos_b_2019.items_authors_institutions                                        
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
    inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions      
    inner join
        rp_author_table                          
            on (
                rp_author_table.fk_items = wos_b_2019.items_authors_institutions.fk_items                  
                and rp_author_table.fk_authors = wos_b_2019.items_authors_institutions.fk_authors             
            )
```

Store in personal table space

```{r}
dbWriteTable(conn = jdbcConnection, 
             name = "rp_countries_table", 
             value = rp_countries)
```

#### Aggregate `rp_countries_table`

Counting method: country of affilation is only counted once per article.

```{sql connection=jdbcConnection}
select
    countrycode,
    issn,
    count(distinct(ut_eid)) as publications 
from
    rp_countries_table 
group by
    countrycode,
    issn
order by issn
```

### 2019-09-08 analysis/002_kb_rp_coverage.Rmd
initial upload

---
title: "Reprint Author Analysis"
output: github_document
bibliography: pubs.bib
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)

options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r setup, echo = FALSE}
# deal with rJava memory allocation 
# https://stackoverflow.com/questions/34624002/r-error-java-lang-outofmemoryerror-java-heap-space
options(java.parameters = "-Xmx1024m")
require(tidyverse)
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar")
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    Sys.getenv("kb_user"),
    Sys.getenv("kb_pwd")
  ) 
```

## Questions

1. How many records in the Web of Science have affiliation information about reprint authors?

2. How many records in the Web of Science have more than one reprint author?

3. How many reprint authors are co-located (institutional level)?

4. How many reprint authors are internationally co-located (country level)?

5. What is the collaboration network in terms of country affiliations from reprint authors?

Focus:

- Database: wos_b_2019
- Document Types: Articles and Reviews
- Database Collections: `WOS.SCI`, `WOS.SSCI`, `WOS.AHCI`
- Publication Period 2014 - 2017

## Data analysis

### 1. How many records in the Web of Science have affiliation information at the country level about reprint authors?

#### Articles total

```{sql, connection=jdbcConnection, output.var="articles_total"}
select wos_b_2019.items.pubyear, COUNT(DISTINCT(wos_b_2019.items.ut_eid)) as articles_total
    from
        wos_b_2019.items      
    inner join
        wos_b_2019.databasecollection              
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items        
    where
        wos_b_2019.databasecollection.edition_value in ('WOS.SCI', 'WOS.SSCI', 'WOS.AHCI')                  
        and wos_b_2019.items.doctype in ('Article', 'Review')                   
        and wos_b_2019.items.pubyear in (2014, 2015, 2016, 2017, 2018)
group by wos_b_2019.items.pubyear
```

```{r}
articles_total %>% 
  arrange(desc(PUBYEAR))
```

#### Records with reprint authors

Using `wos_b_2019.items_authors_institutions.type = 'RP'` to identify reprint authors.

```{sql, connection=jdbcConnection, output.var="rp_articles_total"}
select wos_b_2019.items.pubyear, COUNT(DISTINCT(wos_b_2019.items.ut_eid)) as rp_articles_total
    from
        wos_b_2019.items      
    inner join
        wos_b_2019.databasecollection              
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items        
    inner join
        wos_b_2019.items_authors_institutions                                        
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items     
    where
        wos_b_2019.databasecollection.edition_value in ('WOS.SCI', 'WOS.SSCI', 'WOS.AHCI')                  
        and wos_b_2019.items.doctype in ('Article', 'Review')                   
        and wos_b_2019.items.pubyear in (2014, 2015, 2016, 2017, 2018)              
        and wos_b_2019.items_authors_institutions.type = 'RP'
    group by wos_b_2019.items.pubyear
```

```{r}
rp_articles_total %>% 
  arrange(desc(PUBYEAR))
```

#### Reprint authors records with country information

*Retrieval considerations*

[An earlier exploration](kb_rp_exploration.md) studying how affiliations from reprint authors are represented in the WOS-KB revealed that only one affiliation is tagged with `RP in the table `wos_b_2019.items_author_institutions` per author. To obtain all affiliations per reprint author, the following retrieval strategy can be used:

1. Get all authors tagged as RP per item
2. Obtain all institutional addresses per rp_author-item combination including countrycode
3. Summarize observations

*Country codes*

There are two tables containing country codes, `wos_b_2019.institutions.countrycode` and `wos_b_2019.d_items_authors_institutions.inst_countrycode`. First, *`wos_b_2019.institutions.countrycode`* is analyzed

```{sql, connection=jdbcConnection, output.var="rp_cc_articles_total"}
select
        d.pubyear, count(distinct(d.ut_eid)) as rp_cc_articles_total
    from
        wos_b_2019.items_authors_institutions 
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                pubyear
            from
                wos_b_2019.items     
            inner join
                wos_b_2019.databasecollection                           
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items             
            inner join
                wos_b_2019.items_authors_institutions                                                     
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'
                )                           
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'
                )                            
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018 
                )                       
                and wos_b_2019.items_authors_institutions.type = 'RP'         
        ) d 
            on (
                wos_b_2019.items_authors_institutions.fk_items = d.pk_items 
                and wos_b_2019.items_authors_institutions.fk_authors = d.fk_authors
            )
         inner join wos_b_2019.institutions on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions
         where wos_b_2019.institutions.countrycode is not null     
   group by d.pubyear
```

```{r}
rp_cc_articles_total %>% 
  arrange(desc(PUBYEAR))
```

Data Summary:

```{r}
rq_1_df <- articles_total %>%
  left_join(rp_articles_total, by = "PUBYEAR") %>%
  mutate(ARTICLE_RP_RATIO = RP_ARTICLES_TOTAL / ARTICLES_TOTAL) %>%
  left_join(rp_cc_articles_total, by = "PUBYEAR") %>%
  mutate(ARTICLE_CC_RP_RATIO = RP_CC_ARTICLES_TOTAL / ARTICLES_TOTAL) %>%
  arrange(desc(PUBYEAR))
rq_1_df
```

```{r}
rq_1_df %>%
  gather(ARTICLES_TOTAL, RP_ARTICLES_TOTAL, RP_CC_ARTICLES_TOTAL, 
         key = "indicator", value = "wos_articles") %>%
  ggplot(aes(PUBYEAR, wos_articles, fill = indicator)) +
  geom_bar(position="dodge", stat="identity") +
  scale_fill_viridis_d("", option = "C", labels = c("Articles + Reviews", "with RP", "with RP and countrycode")) +
  theme_minimal() +
  labs(y = "wos_b_2019 records", 
       title = "Coverage of reprint authors with country of affiliation",
       subtitle = "wos_b_2019.institutions.countrycode") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

Next, *`wos_b_2019.d_items_authors_institutions.inst_countrycode`* is analyzed.

```{sql connection=jdbcConnection, output.var="rp_d_cc_articles_total"}
  select
        d.pubyear, count(distinct(d.ut_eid)) as rp_d_cc_articles
    from
        wos_b_2019.d_items_authors_institutions 
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                pubyear
            from
                wos_b_2019.items     
            inner join
                wos_b_2019.databasecollection                           
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items             
            inner join
                wos_b_2019.items_authors_institutions                                                     
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'
                )                           
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'
                )                            
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018
                )                       
                and wos_b_2019.items_authors_institutions.type = 'RP'         
        ) d 
            on (
                wos_b_2019.d_items_authors_institutions.fk_items = d.pk_items 
                and wos_b_2019.d_items_authors_institutions.fk_authors = d.fk_authors
            )
         where wos_b_2019.d_items_authors_institutions.inst_countrycode is not null     
   group by d.pubyear
```

```{r}
rp_d_cc_articles_total
```

```{r}
rp_df_d <- articles_total %>%
  left_join(rp_articles_total, by = "PUBYEAR") %>%
  mutate(ARTICLE_RP_RATIO = RP_ARTICLES_TOTAL / ARTICLES_TOTAL) %>%
  left_join(rp_d_cc_articles_total, by = "PUBYEAR") %>%
  mutate(RP_D_CC_ARTICLES_RATIO = RP_D_CC_ARTICLES / ARTICLES_TOTAL) %>%
  arrange(desc(PUBYEAR))
rp_df_d
```

```{r}
rp_df_d %>%
  gather(ARTICLES_TOTAL, RP_ARTICLES_TOTAL, RP_D_CC_ARTICLES, 
         key = "indicator", value = "wos_articles") %>%
  ggplot(aes(PUBYEAR, wos_articles, fill = indicator)) +
  geom_bar(position="dodge", stat="identity") +
  scale_fill_viridis_d("", option = "C", labels = c("Articles + Reviews", "with RP", "with RP and countrycode")) +
  theme_minimal() +
  labs(y = "wos_b_2019 records", 
       title = "Coverage of reprint authors with country of affiliation",
       subtitle = "wos_b_2019.d_items_authors_institutions.inst_countrycode") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

In conclusion, joining with both tables seems feasible, because it results in a thorough countries of affiliation coverage for reprint authors. Overall, `r round(mean(rp_df_d$RP_D_CC_ARTICLES_RATIO) * 100, 2)` % of records representing original articles and reviews in the period 2014 - 2017 provide affiliation information for reprint authors at the country-level. In the following, `wos_b_2019.d_items_authors_institutions.inst_countrycode` is used.

### 2. How many records in the Web of Science have more than one reprint author?

#### Query

```{sql connection=jdbcConnection, output.var="rp_ind_df"}
select
        distinct         
        d.pubyear,
        d.ut_eid,
        wos_b_2019.d_items_authors_institutions.fk_authors,
        wos_b_2019.d_items_authors_institutions.inst_per_auth_cnt,
        wos_b_2019.d_items_authors_institutions.country_per_auth_cnt     
    from
        wos_b_2019.d_items_authors_institutions      
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                pubyear             
            from
                wos_b_2019.items                  
            inner join
                wos_b_2019.databasecollection                                                
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items                          
            inner join
                wos_b_2019.items_authors_institutions                                                                          
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items                               
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'                 
                )                                            
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'                 
                )                                             
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018                 
                )                                        
                and wos_b_2019.items_authors_institutions.type = 'RP'                  
        ) d              
            on (
                wos_b_2019.d_items_authors_institutions.fk_items = d.pk_items                  
                and wos_b_2019.d_items_authors_institutions.fk_authors = d.fk_authors             
            )          
    where
        wos_b_2019.d_items_authors_institutions.inst_countrycode is not null
```

#### Backup in KB table space

```{r}
dbWriteTable(conn = jdbcConnection, 
             name = "rp_ind_df", 
             value = rp_ind_df,
             overwrite = TRUE)
```

#### Analysis: Distribution of the number of reprint authors

```{sql connection=jdbcConnection, output.var="rp_au"}
select
        pubyear,
        ut_eid,
        count(fk_authors) as n_au
    from
        rp_ind_df 
    group by
        ut_eid,
        pubyear 
```

Summary statistics

```{r}
summary(rp_au$N_AU)
```

Frequency table of the number of reprint authors per publication. Publications with more than five reprint authors per publication were reduced to the residual category "> 5".

```{r}
tp_au_count <- rp_au %>% 
  mutate(N_AU = fct_other(factor(N_AU), keep = c("1", "2", "3", "4", "5"),
                          other_level = "> 5")) %>% 
  group_by(N_AU, PUBYEAR) %>% 
  summarize(N = n()) %>% 
  ungroup() %>% 
  group_by(PUBYEAR) %>% 
  mutate(PROP = N / sum(N))
tp_au_count %>%
  ggplot(aes(N_AU, PROP, fill = factor(PUBYEAR))) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  facet_grid( ~ factor(PUBYEAR)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  labs(
    x = "Number of reprint authors per record",
    y = "Proportion",
    title = "Number of reprint authors per Web of Science record",
    subtitle = "Articles and reviews 2014 - 2018"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

SQL statement to retrieve all publications with more than one reprint author (incl. country info)

```{sql connection=jdbcConnection, output.var="rp_au_count"}
select
        pubyear,
        ut_eid,
        count(fk_authors)  
    from
        rp_ind_df 
    group by
        ut_eid,
        pubyear 
    having
        count(fk_authors) > 1
```

In total, `r length(unique(rp_au_count$UT_EID))` out of `r length(unique(rp_ind_df$UT_EID))` records indexed in the Web of Science for the period 2014 - 2017 have more than one reprint author, representing a percentage of `r round(length(unique(rp_au_count$UT_EID)) / length(unique(rp_ind_df$UT_EID)) * 100, 2)` %. Furthermore, data exploration suggests that starting from 2016 the Web of Science has begun to systematically keep track of more than one reprint author.

### 3. How many reprint authors are co-located (institutional level)?

#### Query

```{sql connection=jdbcConnection, output.var="rp_inst_au_count"}
select
        inst_per_auth_cnt,
        pubyear,
        count(*) as inst_n
    from
        rp_ind_df 
    group by
        pubyear,
        inst_per_auth_cnt
```

#### Analysis

Distribution of the number of institutional affiliations per reprint author and year

```{r}
rp_inst_au_df <- rp_inst_au_count %>%
  mutate(INST_PER_AUTH_CNT = fct_other(
    factor(INST_PER_AUTH_CNT),
    keep = c("1", "2", "3", "4", "5"),
    other_level = "> 5"
  )) %>%
  ungroup() %>%
  group_by(PUBYEAR) %>%
  mutate(PROP = INST_N / sum(INST_N)) %>%
  arrange(INST_PER_AUTH_CNT, PUBYEAR)

rp_inst_au_df %>%
  ggplot(aes(
    x = factor(INST_PER_AUTH_CNT),
    PROP,
    fill = factor(PUBYEAR)
  )) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  facet_grid( ~ PUBYEAR) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  labs(
    x = "Number of Institutions",
    y = "Proportion",
    title = "Institutional affiliations per reprint author",
    caption = "Derived from wos_b_2019.D_ITEMS_AUTHORS_INSTITUTIONS.INST_PER_AUTH_CNT"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

Overall, around half of the reprint authors indexed listed more than one institutional affiliation (N = `r rp_inst_au_count %>% filter(!INST_PER_AUTH_CNT == "1") %>% .$INST_N %>% sum()`, Proportion = `r round(rp_inst_au_count %>% filter(!INST_PER_AUTH_CNT == "1") %>% .$INST_N %>% sum() / rp_inst_au_count  %>% .$INST_N %>% sum(), 4) * 100` %).

### 4. How many reprint authors are internationally co-located (country level)?

#### Query

```{sql connection=jdbcConnection, output.var="rp_country_au_count"}
select
        country_per_auth_cnt,
        pubyear,
        count(*) as country_n
    from
        rp_ind_df 
    group by
        pubyear,
        country_per_auth_cnt
```

Frequency table of the number of country affiliation per reprint author and publication. Reprint authors with more than three countries of affiliation at the time of publication were reduced to the residual category "Other".

```{r}
rp_country_au_df <- rp_country_au_count %>%
  mutate(COUNTRY_PER_AUTH_CNT = fct_other(
    factor(COUNTRY_PER_AUTH_CNT),
    keep = c("1", "2", "3"),
    other_level = "> 3"
  )) %>%
  ungroup() %>%
  group_by(PUBYEAR) %>%
  mutate(PROP = COUNTRY_N / sum(COUNTRY_N)) %>%
  arrange(COUNTRY_PER_AUTH_CNT, PUBYEAR)
```

```{r}
rp_country_au_df
```

```{r}
rp_country_au_df %>%
  ggplot(aes(
    x = factor(COUNTRY_PER_AUTH_CNT),
    PROP,
    fill = factor(PUBYEAR)
  )) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  facet_grid( ~ PUBYEAR) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  labs(
    x = "Number of Countries",
    y = "Proportion",
    title = "Country affiliations per reprint author",
    caption = "Derived from wos_b_2019.D_ITEMS_AUTHORS_INSTITUTIONS.COUNTRY_PER_AUTH_CNT"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

Although the number of co-located reprint authors is growing, only a small proportion of reprint author is internationally co-located (N = `r rp_country_au_df %>% filter(!COUNTRY_PER_AUTH_CNT == "1") %>% .$COUNTRY_N  %>% sum()`, Proportion = `r round(rp_country_au_df %>% filter(!COUNTRY_PER_AUTH_CNT == "1") %>% .$COUNTRY_N  %>% sum() / rp_country_au_df  %>% .$COUNTRY_N  %>% sum(), 4) * 100` %).

### 5. What is the collaboration network in terms of country affiliations from reprint authors?

#### Query

Obtain publications with country affiliations

```{sql connection=jdbcConnection, output.var="rp_country_mat"}
select  distinct wos_b_2019.d_items_authors_institutions.INST_COUNTRYCODE,       
        d.ut_eid
    from
        wos_b_2019.d_items_authors_institutions      
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                pubyear             
            from
                wos_b_2019.items                  
            inner join
                wos_b_2019.databasecollection                                                
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items                          
            inner join
                wos_b_2019.items_authors_institutions                                                                          
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items                               
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'                 
                )                                            
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'                 
                )                                             
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018                 
                )                                        
                and wos_b_2019.items_authors_institutions.type = 'RP'                  
        ) d              
            on (
                wos_b_2019.d_items_authors_institutions.fk_items = d.pk_items                  
                and wos_b_2019.d_items_authors_institutions.fk_authors = d.fk_authors             
            )          
    where
        wos_b_2019.d_items_authors_institutions.inst_countrycode is not null
```

#### Analysis

Publications with internationally co-located reprint author(s)

```{r}
co_pubs <- rp_country_mat %>% 
  group_by(UT_EID) %>%
  filter(n() > 1)
```

Country share

```{r}
all_count <- rp_country_mat %>%
  count(INST_COUNTRYCODE, sort = TRUE)
co_count <- co_pubs %>%
  ungroup() %>%
  count(INST_COUNTRYCODE, sort = TRUE, name = "n_co")
left_join(all_count, co_count, by = "INST_COUNTRYCODE") %>%
  mutate(prop = n_co / n * 100) 
```

Create bi-partite matrix

```{r}
#co_pubs_mat <- as.matrix(table(co_pubs$UT_EID, co_pubs$INST_COUNTRYCODE))
```

calculate unipartite matrix (country level)

```{r}
#co_pubs_mat_t <- t(co_pubs_mat) %*% co_pubs_mat
```

visualise network (tbc)

```{r}
# library(sna)
# sna::gplot(co_pubs_mat_t, 
#            gmode = "graph", 
#            vertex.cex = log(diag(co_pubs_mat_t)^0.3),
#            usearrows = FALSE)
```

### 2019-09-08 analysis/003_kb_fetch_ca.Rmd
initial upload

---
title: "Fetching Reprint Author info - v2"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

## About

Country information about corresponding authors play a crucial role in open access funding (Schimmer, Geschuhn, and Vogler 2015). The Web of Science contains data about this author role including affiliations, making it a valuable source to determine publisher's market shares per country in terms of articles published.

In this work, a dataset is compiled providing information about productivity of corresponding authors between 2014 - 2018. More specificially, the following data points from the Web of of Science will be investigated:

- Web of Science collections SCI, SSCI and WOS.AHCI 
- Article types Original Articles and Reviews
- Country of affiliations
- ISSN
- Publication year

This dataset will be extended with:

- Publisher and journal titles from Crossref
- Journal open access status provided by the Bielefeld GOLD OA list

In this document, every step needed to create the dataset is described. This document is written in [RMarkdown](https://rmarkdown.rstudio.com/). It can be re-compiled calling

```r
require(rmarkdown)
require(tidyverse)
require(RJDBC)
require(rJava)
require(writexl)

rmarkdown::render("kb_fetch_new.Rmd")
```

## Workflow

### Obtain publication data from the Web of Science database hosted by the German Competence Centre for Bibliometrics

#### Connect to database

```{r}
# deal with rJava memory allocation 
# https://stackoverflow.com/questions/34624002/r-error-java-lang-outofmemoryerror-java-heap-space
options(java.parameters = "-Xmx1024m")
require(tidyverse)
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar")
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    Sys.getenv("kb_user"),
    Sys.getenv("kb_pwd")
  ) 
```

#### Query

```{sql connection=jdbcConnection, output.var="rp_country"}
select
        wos_b_2019.d_items_authors_institutions.inst_countrycode,
        d.pubyear,
        d.issn,
        count(distinct(d.ut_eid)) as pubs
    from
        wos_b_2019.d_items_authors_institutions           
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                wos_b_2019.items.pubyear,
                issn             
            from
                wos_b_2019.items                               
            inner join
                wos_b_2019.databasecollection                                                                     
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items               
            inner join
                wos_b_2019.items_authors_institutions                                     
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items             
            inner join
                wos_b_2019.issues                                   
                    on wos_b_2019.issues.pk_issues = wos_b_2019.items.fk_issues                
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'                                  
                )                                                             
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'                                  
                )                                                              
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018                                  
                )                                                         
                and wos_b_2019.items_authors_institutions.type = 'RP'                           
        ) d                           
            on (
                wos_b_2019.d_items_authors_institutions.fk_items = d.pk_items                                   
                and wos_b_2019.d_items_authors_institutions.fk_authors = d.fk_authors                          
            )               
    where
        wos_b_2019.d_items_authors_institutions.inst_countrycode is not null     
    group by
        wos_b_2019.d_items_authors_institutions.inst_countrycode,
        d.pubyear,
        d.issn
```

```{r}
# dump
write_csv(rp_country, "data/rp_data_14_18.csv")
```

### Get most frequent publisher and journal names from Crossref

Publisher names used in the Web of Science are ambigue, and sometimes journals can change the publisher. To obtain most frequent publisher and journals names, Crossref was queried for every distinct journal by ISSN as described in [`fetch_data.R`](fetch_data.R).

As a result, a [matching table between Crossref and the Web of Science journals](data/cr_wos.csv) was created. In the following, this table is joined with the WoS data.

```{r}
cr_journals <- readr::read_csv("data/cr_wos.csv") %>%
  # merge springer nature brands
  mutate(publisher = ifelse(grepl("Springer", publisher, fixed = FALSE), "Springer Nature", publisher))
rp_country <- readr::read_csv("data/rp_data_14_18.csv")
issn_l <- readr::read_tsv("data/20190818.ISSN-to-ISSN-L.txt") %>%
  # manual fix Journal - American Water Works 
  add_row(ISSN = "2164-4535", `ISSN-L` = "0003-150X")
rp_df <- rp_country %>%
  left_join(issn_l, by = "ISSN") %>%
  left_join(cr_journals, by = c(`ISSN-L`= "issn_l")) %>%
  select(-issn) %>%
  distinct()
```

Crossref indexed `r rp_df %>% filter(!is.na(publisher)) %>% distinct(ISSN) %>% nrow() / rp_df %>% distinct(ISSN) %>% nrow() * 100` % of Web of Science journals.

The following table shows the number of journals per publisher.

```{r}
rp_df %>% 
  distinct(ISSN, publisher) %>% 
  count(publisher, sort = TRUE) %>% 
  mutate(prop = n /sum(n))
```

### 3. Obtain open access status information

Bruns et al. provide a matching table for fully open access journals, joining multiple sources into one dataset.

Bruns, A., Lenke, C., Schmidt, C., & Taubert, N. C. (2019). ISSN-Matching of Gold OA Journals (ISSN-GOLD-OA) 3.0. Bielefeld University. doi:[10.4119/unibi/2934907](https://doi.org/10.4119/unibi/2934907)

The linking ISSN is used for matching.

```{r}
u <- "https://pub.uni-bielefeld.de/download/2934907/2934908/ISSN_Gold-OA_3.0.csv"
bie_oa <- readr::read_csv(u) %>%
  # remove missing entries with missing ISSN_L
  filter(!is.na(ISSN_L))
# add info to rp_df
rp_df_oa <- rp_df %>%
  mutate(oa_journal = `ISSN-L` %in% bie_oa$`ISSN_L`)
```

The following table shows the number and proportion of open access journals

```{r}
rp_df_oa %>% 
  distinct(`ISSN-L`, oa_journal) %>% 
  count(oa_journal) %>%
  mutate(prop = n / sum(n))
```

### 4. Data coding and dump

`CSV`-based dump

```{r}
rp_df_oa %>%
  select(
    issn_wos = ISSN,
    country_code = INST_COUNTRYCODE,
    publication_year = PUBYEAR,
    articles = PUBS,
    journal_title, publisher, oa_journal,
    issn_l = `ISSN-L`) %>%
  write_csv("data/rp_jn_14_18.csv")
```

Excel dump. To avoid loading issues because of too many rows in one sheet, split the dataset into several sheets per year.

```{r}
require(writexl)
rp_df <- readr::read_csv("data/rp_jn_14_18.csv")
tmp <- map(2014:2018, function(x) {
  rp_df %>% 
    filter(publication_year == x)
  }
)
# names
names(tmp) <- paste0("pubyear_", 2014:2018)
write_xlsx(tmp, "data/rp_14_18.xlsx")
```

Data Schema:

|Variable           |Description                                                                                                                                                                      |Source                                                                                                                                                       |
|:------------------|:--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|:------------------------------------------------------------------------------------------------------------------------------------------------------------|
|`issn_wos`         |ISSN, a standardized journal id.                                                                                                                                                 |KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                  |
|`country_code`     |Country of affiliation corresponding author, represented as ISO 3 code                                                                                                           |KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode`                                                                                |
|`publication_year` |Year of publication, obtained from KB Web of Science                                                                                                                             |KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                                |
|`articles`         |Number of original articles and reviews published. Whole counting where internationally co-located corresponding authorships were assigned to each contributing country equally. |KB Web of Science: Grouped counts over `wos_b_2019.issues.issn`,  `wos_b_2019.d_items_authors_institutions.inst_countrycode` and  `wos_b_2019.items.pubyear` |
|`journal_title`    |Most frequently used journal title in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                       |Crossref                                                                                                                                                     |
|`publisher`        |Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                      |Crossref                                                                                                                                                     |
|`oa_journal`       |Is the journal publishing all articles open access without delay (full open access)?                                                                                             |Bielefeld GOLD OA List V3                                                                                                                                    |
|`issn_l`           |Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.                                             |CIEPS                                                                                                                                                        |

### 2019-09-08 analysis/004_kb_fetch_publisher.Rmd
initial upload

--- 2019-09-08 analysis/003_kb_fetch_ca.Rmd
@@
 ---
-title: "Fetching Reprint Author info - v2"
+title: "Publisher League Global"
+author: "Najko Jahn"
+date: "9/5/2019"
 output: github_document
@@
 
-Country information about corresponding authors play a crucial role in open access funding (Schimmer, Geschuhn, and Vogler 2015).
-The Web of Science contains data about this author role including affiliations, making it a valuable source to determine publisher's market shares per country in terms of articles published.
-
-In this work, a dataset is compiled providing information about productivity of corresponding authors between 2014 - 2018.
-More specificially, the following data points from the Web of of Science will be investigated:
-
-- Web of Science collections SCI, SSCI and WOS.AHCI 
-- Article types Original Articles and Reviews
-- Country of affiliations
-- ISSN
-- Publication year
-
-This dataset will be extended with:
-
-- Publisher and journal titles from Crossref
-- Journal open access status provided by the Bielefeld GOLD OA list
-
-In this document, every step needed to create the dataset is described. This document is written in [RMarkdown](https://rmarkdown.rstudio.com/). It can be re-compiled calling
+This executable report describes how data about the global market shares of publishers were retrieved from the Web of Science (WoS) in-house database hosted by the [German Competence Center for Bibliometrics (KB)](http://www.bibliometrie.info/).
+This document is written in R Markdown.
+If you have access to WoS-KB data infrastructure, you will be able to replicate the methods used by calling
 
@@
 require(writexl)
-
-rmarkdown::render("kb_fetch_new.Rmd")
+rmarkdown::render("kb_fetch_publisher_global.Rmd")
 ```
@@
 jdbcDriver <-
-  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar")
+  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar") # you may need to change the path to your db driver
 jdbcConnection <-
@@
     "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
+    # login credentials are stored in the .Rprofile file 
     Sys.getenv("kb_user"), 
@@
 
-```{sql connection=jdbcConnection, output.var="rp_country"}
+```{sql connection=jdbcConnection, output.var="rp_publisher"}
 select
-        wos_b_2019.d_items_authors_institutions.inst_countrycode,
-        d.pubyear,
-        d.issn,
-        count(distinct(d.ut_eid)) as pubs
-    from
-        wos_b_2019.d_items_authors_institutions           
-    inner join
-        (
-            select
-                pk_items,
-                fk_authors,
-                ut_eid,
         wos_b_2019.items.pubyear,
-                issn             
+        wos_b_2019.issues.issn,
+        count(distinct(ut_eid)) as pubs                         
     from
@@
             on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items                                  
-            inner join
-                wos_b_2019.items_authors_institutions                                     
-                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items             
     inner join
@@
         )                                                  
-                and wos_b_2019.items_authors_institutions.type = 'RP'                           
-        ) d                           
-            on (
-                wos_b_2019.d_items_authors_institutions.fk_items = d.pk_items                                   
-                and wos_b_2019.d_items_authors_institutions.fk_authors = d.fk_authors                          
-            )               
-    where
-        wos_b_2019.d_items_authors_institutions.inst_countrycode is not null     
     group by
-        wos_b_2019.d_items_authors_institutions.inst_countrycode,
-        d.pubyear,
-        d.issn
+        wos_b_2019.items.pubyear,
+        wos_b_2019.issues.issn
 ```
@@
 ```{r}
+head(rp_publisher) %>%
+  knitr::kable()
 # dump
-write_csv(rp_country, "data/rp_data_14_18.csv")
+write_csv(rp_publisher, "data/publisher_league_14_18.csv")
 ```
@@
   mutate(publisher = ifelse(grepl("Springer", publisher, fixed = FALSE), "Springer Nature", publisher))
-rp_country <- readr::read_csv("data/rp_data_14_18.csv")
+rp_publisher <- readr::read_csv("data/publisher_league_14_18.csv")
 issn_l <- readr::read_tsv("data/20190818.ISSN-to-ISSN-L.txt") %>%
@@
   add_row(ISSN = "2164-4535", `ISSN-L` = "0003-150X")
-rp_df <- rp_country %>%
+rp_df <- rp_publisher %>%
   left_join(issn_l, by = "ISSN") %>%
@@
   count(publisher, sort = TRUE) %>% 
-  mutate(prop = n /sum(n))
+  mutate(prop = n /sum(n) * 100)
+```
+
+Next, the number and proportion of original articles and reviews published per publisher is presented:
+
+```{r}
+rp_df %>% 
+  group_by(publisher) %>%
+  summarise(n = sum(PUBS)) %>% 
+  mutate(prop = n /sum(n) * 100) %>%
+  arrange(desc(prop))
 ```
@@
     issn_wos = ISSN,
-    country_code = INST_COUNTRYCODE,
     publication_year = PUBYEAR,
@@
     journal_title, publisher, oa_journal,
-    issn_l = `ISSN-L`) %>%
-  write_csv("data/rp_jn_14_18.csv")
+    issn_l = `ISSN-L`
+    ) %>%
+  write_csv("data/publisher_14_18.csv")
 ```
 
-Excel dump. To avoid loading issues because of too many rows in one sheet, split the dataset into several sheets per year.
+Dump to Excel
 
 ```{r}
-require(writexl)
-rp_df <- readr::read_csv("data/rp_jn_14_18.csv")
-tmp <- map(2014:2018, function(x) {
-  rp_df %>% 
-    filter(publication_year == x)
-  }
-)
-# names
-names(tmp) <- paste0("pubyear_", 2014:2018)
-write_xlsx(tmp, "data/rp_14_18.xlsx")
+readr::read_csv("data/publisher_14_18.csv") %>%
+  writexl::write_xlsx("data/publisher_14_18.xlsx")
 ```
+
+Data Schema:
 
@@
 |KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                  |
-|`country_code`     |Country of affiliation corresponding author, represented as ISO 3 code                                                                                                           |KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode`                                                                                |
 |`publication_year` |Year of publication, obtained from KB Web of Science                                                                                                                             |KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                                |
-|`articles`         |Number of original articles and reviews published.
-Whole counting where internationally co-located corresponding authorships were assigned to each contributing country equally.
-|KB Web of Science: Grouped counts over `wos_b_2019.issues.issn`,  `wos_b_2019.d_items_authors_institutions.inst_countrycode` and  `wos_b_2019.items.pubyear` |
+|`articles`         |Number of original articles and reviews published. |KB Web of Science: Grouped counts over `wos_b_2019.issues.issn` and  `wos_b_2019.items.pubyear` |
 |`journal_title`    |Most frequently used journal title in terms of articles published between 2014 - 2018.

### 2019-09-08 analysis/paper.Rmd
initial upload

---
title: "Data Descriptive: Corresponding author journal publishing activities 2014 - 2018"
date: "9/5/2019"
output: github_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)

options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
library(tidyverse)
```

## Background & summary

The volume of corresponding author publications is a crucial indicator in the planning of open access funding programs and licensing models (Schimmer, Geschuhn, and Vogler 2015). But funders and national library consortia often lack access to such bibliometric data. Particularly, determining the share per publisher, journal and journal business model across countries is challenging. Here, by obtaining country affiliations from the Web of Science at the level of journals, and combining them with normalized publisher and open access status information, we created a comprehensive dataset about the productivity of corresponding authors for the period 2014-2018. This dataset is expected to have essential value for the quantitatve understanding of scholarly publishing. It will not only help to analyze the market share of publishing houses, but will also uncover the current state of the transition of subscription-based publishing models to fully open access.

Only a few openly available datasets focusing on the scholarly publishing industry exist. For example, Larivière et al. (2015), shared an aggregated dataset that contains the market share of publishers in terms of articles indexed in the Web of Science between 1973-2013. Using this data, the authors were able to reveal a growing market concentration among a few publishing houses. Palzenberger (2015) provided a dataset presenting the number and proportion of scholarly articles per country of affiliation including a breakdown per author role for the period 2004-2014. Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to open access business model is feasible. Since then, analyzing the productivity of corresponding authors in terms of articles published in dedicated journal portfolios has been perceived as a critical component of open access negotiations.

There have also been analytical applications for such datasets. Major bibliometric databases including the Web of Science, Scopus and Dimension have started to integrate open access status information. In these databases, Unpaywall data is often used to identify openly available full-texts at the article-level, while the Directory of Open Access Journals provides information about whether the journal is operating under a fully open access or toll-access business model. In its most recent edition, the Leiden Ranking presents open access indicators at the level of universities. In addition to university-specific data, the German Open Access Monitor, provides breakdowns per publisher and journal, as well as the inclusion of publications from non-university research institutions based in Germany

Here, we build an open dataset on the global productivity of corresponding authors. Drawing on Palzenberger's previous work, we not only created an updated dataset about the number and proportion of scholarly articles per country of affiliation from corresponding authors. The dataset also allows for a breakdown by journal, including information whether the journal is currently fully open access or not, and publishers. We constructed the dataset using the quality assured Web of Science in-house database from the German Competence Center for Bibliometrics. We enriched the data with publisher information from Crossref, a large DOI registration agency for scholarly work, and with open access journal information provided by the Biefeld GOLD OA list.

The data cannot only serve as a critical input for negotiations in the context of transformative agreements, but also in the quantitative study of the academic publishing industry and the open access uptake across countries. In the latter case, for instance, national transformative agreements with large publishers will likely have an impact on the open access share of a country. To demonstrate its potential significance, we will present two new findings using the dataset: a global map of publishers's toll-access market share, and an analysis of country of affiliation per publisher. Source code used for compiling the dataset is shared along with the resulting datasets.

## Data and Methods

Following Palzenberger, data were retrieved from the in-house Web of Science database maintained by the German Competence Center for Bibliometrics. The following data were obtained and aggregated:

- Web of Science collections SCI, SSCI and AHCI
- Article types Original Articles and Reviews
- Country of affiliations
- ISSN
- Publication years 2014 - 2018 
- Author Roles

Our database retrieval reflected that it has become challenging to count corresponding authors straight, because the Web of Science has begun to keep track of more than one reprint author recently. While evaluating an efficient retrieval strategy, we found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018. Moreover, a percentage of 5% of reprint author was internationally co-located (see Figure 1). We therefore decided to use whole counting where every Web of Science record representing an original article or review was counted once per country of affiliation.

![](figure/country_wos.png)

After obtaining aggregated article counts at the country and journal level form the Web of Science, Crossref’s REST API was queried to retrieve publisher and journal titles. To control developments of the publishing market resulting in name changes of publishers or journal titles over time, only the most frequent facet field name was used. Drawing on ISSN-L, ISSN variants were obtained to improve the retrieval. Next, the open access status per journal was added to the dataset. This information was retrieved from Bielefeld, which joins open access journal information from various sources including the DOAJ, PubMed Central and the Directory of Open Access Scholarly Resources (ROAD). Again, ISSN-L were used for the matching procedure.

The underlying source code in R and SQL for both testing the retrieval and compiling the datasets are openly shared as executable reports, making our data work more reproducible and transparent.

## Data Records

Overall, our dataset consists of two separate rectangular data, which are openly available at XXXX. Each data is provided in comma-separated values (csv) files and as Excel spreadsheets.

### Global publisher data

```{r}
publisher_df <- readr::read_csv("data/publisher_14_18.csv")
```

The first file, named `publisher_14_18`, contains information about the publishing market landscape for the period 2014-2018 in terms of articles and reviews indexed in the Web of Science. Table 1 describes data variables in detail. We were able to retrieve publisher and journal titles from Crossref for `r publisher_df %>% filter(!is.na(publisher)) %>% distinct(issn_l) %>% nrow() / publisher_df %>% distinct(issn_l) %>% nrow() * 100` %
 of journals, representing `r publisher_df %>% filter(!is.na(publisher)) %>% .$articles %>% sum() / sum(publisher_df$articles) * 100` % of the overall article volume. `r publisher_df %>% filter(oa_journal == TRUE) %>% distinct(issn_l) %>% nrow()` out of `r publisher_df %>% filter(!is.na(publisher)) %>% distinct(issn_l) %>% nrow()`
 investigated journals with `ISSN-L` were identified as fully open access journals. In total, the dataset provides information for `r sum(publisher_df$articles)` records indexed in the Web of Science with journal information in the five-years period 2014-2018.

Data Schema `publisher_14_18`:

|| Variable           | Description                                                                                                                                 | Source                                                                                         |
| :----------------- | :------------------------------------------------------------------------------------------------------------------------------------------ | :--------------------------------------------------------------------------------------------- |
| `issn_wos`         | ISSN, a standardized journal id.                                                                                                            | KB Web of Science: `wos_b_2019.issues.issn`                                                    |
| `publication_year` | Year of publication, obtained from KB Web of Science                                                                                        | KB Web of Science: `wos_b_2019.items.pubyear`                                                  |
| `articles`         | Number of original articles and reviews published.                                                                                          | KB Web of Science: Grouped counts over `wos_b_2019.issues.issn` and `wos_b_2019.items.pubyear` |
| `journal_title`    | Most frequently used journal title in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref  | Crossref                                                                                       |
| `publisher`        | Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref | Crossref                                                                                       |
| `oa_journal`       | Is the journal publishing all articles open access without delay (full open access)?                                                        | Bielefeld GOLD OA List V3                                                                      |
| `issn_l`           | Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.        | CIEPS                                                                                          |
                                                                                             |

### Corresponding author country-level publisher data

```{r}
rp_df <- readr::read_csv("data/rp_jn_14_18.csv")
```

The file named `rp_14_18` provides a breakdown of publisher and journal information per country of affiliation. Only corresponding authorships, indicated by the role "reprint author" in the Web of Science database, were taken into consideration. Table 1 describes data variables in detail. Overall, 99 % of records representing original articles and reviews in the period 2014 - 2017 provide affiliation information for reprint authors at the country-level.

Data
Schema:

| Variable           | Description                                                                                                                                                                      | Source                                                                                                                                                     |
| :----------------- | :------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- | :--------------------------------------------------------------------------------------------------------------------------------------------------------- |
| `issn_wos`         | ISSN, a standardized journal id.                                                                                                                                                 | KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                |
| `country_code`     | Country of affiliation corresponding author, represented as ISO 3 code                                                                                                           | KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode`                                                                              |
| `publication_year` | Year of publication, obtained from KB Web of Science                                                                                                                             | KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                              |
| `articles`         | Number of original articles and reviews published. Whole counting where internationally co-located corresponding authorships were assigned to each contributing country equally. | KB Web of Science: Grouped counts over `wos_b_2019.issues.issn`, `wos_b_2019.d_items_authors_institutions.inst_countrycode` and `wos_b_2019.items.pubyear` |
| `journal_title`    | Most frequently used journal title in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                       | Crossref                                                                                                                                                   |
| `publisher`        | Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                      | Crossref                                                                                                                                                   |
| `oa_journal`       | Is the journal publishing all articles open access without delay (full open access)?                                                                                             | Bielefeld GOLD OA List V3                                                                                                                                  |
| `issn_l`           | Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.                                             | CIEPS                                                                                                                                                      |

<!--### Corresponding author co-authorship network -->

## Use-cases

### Global publisher market share

Using the data record `publisher_14_18` the global market share of publishing houses in terms of articles and reviews indexed in the Web of Science was analyzed for the period 2014-2018. Table 1 shows the overall market share for the ten largest publishers. Figure 1 presents a breakdown by journal business model, differentiating between toll access and fully open access.

```{r, results='asis'}
my_df <- readr::read_csv("data/publisher_14_18.csv")
tt_all <- my_df %>% 
  group_by(publisher) %>%
  summarise(n = sum(articles)) %>%
  mutate(prop = n /sum(n))  %>%
  mutate(publisher = ifelse(is.na(publisher), "Other", publisher)) %>%
  arrange(desc(n)) %>%
  mutate(publisher_group = fct_other(publisher, keep = publisher[1:11])) %>%
  ungroup() %>%
  group_by(publisher_group) %>%
  summarise(n = sum(n), prop = sum(prop)) %>%
  mutate(publisher_group = fct_reorder(publisher_group, n,)) %>%
  mutate(publisher_group = fct_relevel(publisher_group, "Other")) %>%
  arrange(desc(publisher_group))
knitr::kable(tt_all, caption = "Publisher Market Share 2014 - 2018 in terms of articles and reviews indexed in the Web of Science")
```

```{r}
tt_oa <- my_df %>%
  mutate(publisher_group = ifelse(publisher %in% tt_all$publisher_group, publisher, "Other")) %>%
  group_by(publisher_group, oa_journal) %>%
  summarise(n = sum(articles)) %>%
  mutate(oa_journal = ifelse(oa_journal == TRUE, "Open Access", "Toll Access")) %>%
  ungroup() %>%
  group_by(oa_journal) %>%
  mutate(prop = n / sum(n))
plot_df <- tt_all %>%
  mutate(oa_journal = "All") %>%
  bind_rows(tt_oa) %>%
  group_by(oa_journal) %>% 
  mutate(label = format(sum(n), big.mark   = ",")) %>% 
  mutate(label = paste0(oa_journal, " (", label, " Articles)")) %>%
  ungroup() %>%
  mutate(publisher_group = factor(publisher_group, levels = levels(tt_all$publisher_group)))
tableau_palette <- rev(c("#ccc94d","#82853b","#86b4a9","#39737c","#ffd94a","#b85a0d","#98d9e4","#3cb7cc","#ffb977","#ff7f0f","#acd98d","#32a251"))
p_bar <- ggplot(plot_df, aes(x="", y = prop, fill = publisher_group)) +
  geom_bar(width = 1, stat = "identity", position = position_stack(reverse = FALSE)) +
  facet_wrap(~label, ncol = 1) +
  coord_flip() +
  geom_text(aes(label = paste0(round(prop, 2) * 100, "")), position = position_stack(vjust = .5)) +
 scale_fill_manual("", values = c("grey80", tableau_palette[1:10]), guide = guide_legend(reverse = TRUE,  ncol= 4)) +
  scale_y_continuous(breaks = seq(0,1, by =0.1), labels = scales::percent_format(accuracy = 5L)) +
  labs(title = "Global Publisher Market Shares 2014 - 2018", x = "", y = "",
       subtitle = "Original Articles and Reviews indexed in the Web of Science\n\n",
       caption = "\nData Source:  Web of Science / Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata. OA journals identified using the Bielefeld Gold OA list.\nFunding: Federal Ministry of Education and Research of Germany (BMBF)") +
  #  scale_y_continuous(breaks = seq(0,1, by =0.1), labels = scales::percent_format(accuracy = 5L)) +
  theme_minimal(base_family = "Roboto", base_size = 16) +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.y = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(legend.position="bottom",
        legend.justification = "center")
ggsave(
  filename = "figure/global_publisher_share.png",
  plot = p_bar,
  width = 14,
  height = 9,
  dpi = 300)
```

![](figure/global_publisher_share.png)

Our data confirms that the global publishing landscape is dominated by a few large publishers. In total, ten publishers accounted for 65% of articles published in the period 2018. The Big 3, Elsevier, Springer Nature and Wiley, even accounted for a percentage of 46 %. However, analysis suggests large differences when differentiating between ftoll and fully open access: The market for fully open access journals is generally less domintated by major publishing houses. A notable exeption is Springer Nature with its large fully open access portfolio including BioMed Central or the large mega-journals Scientific Reports and Nature Communications.

### Global Map of Toll-Access Journal Publishing

Using `rp_14_18` the publisher market shares can be furthermore broken down per country of affiliation from corresponding author. Figure presents a global map of publishing in toll-access journals. Publications from institutions based in the European Union were summarized to "EU28". Together, EU 28 member states constituted the largest market for toll-access publishing between 2014-18.

![](figure/worldmap.png)

## Responsible use

Although the Web of Science database is a well-established bibliometric database in the quantitative science studies it comes with certain limitations that needs to considered when interpreting our data. First, it consists of a limited set of journals, likely not covering the whole journal portfolio of the investigated publishers. Although publisher names were disambiguated, journal title transfers between publishers were not tracked. Furthermore, the year of inclusion in a journal issue, i.e. print publication date, was used. However, there can be considerable time-lag from online-first publication and issue inclusion.

As any other bibliometric databases, the data structure can be subject of change. In particular, our pretests suggest that the reprint author field used to identify corresponding authorships has undergone a change recently, starting to covering more than one reprint author per article.

The Bielefeld GOLD OA list was used to identify fully open access journals. One of its weakness is that it does not contain the date of journal flipping. The figures should be therefore interpreted as numbers of articles available in current fully open access journals, instead of published in an fully open access journal at the time of publication.

Last, but not least, qualitative evaluation is still needed. All code is shared to ensure the technical reproducibility of all steps involved in our data work.

## Acknowledgment

## References

### 2019-09-09 analysis/001_kb_rp_pretest.Rmd
rmd reports render

@@
 jdbcDriver <-
-  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar")
+  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "../inst/jdbc_driver/ojdbc8.jar")
 jdbcConnection <-

### 2019-09-09 analysis/002_kb_rp_coverage.Rmd
rmd reports render

@@
 - Database Collections: `WOS.SCI`, `WOS.SSCI`, `WOS.AHCI`
-- Publication Period 2014 - 2017
+- Publication Period 2014 - 2018
 
@@
 
-```{r}
+```{r ca_author_count}
 tp_au_count <- rp_au %>% 
@@
 
-```{r}
+```{r ca_country_count}
 rp_country_au_df %>%

### 2019-09-09 analysis/003_kb_fetch_ca.Rmd
rmd reports render

@@
 ---
-title: "Fetching Reprint Author info - v2"
+title: "Fetching Reprint Author information from the Web of Science"
 output: github_document
@@
 
-rmarkdown::render("kb_fetch_new.Rmd")
+rmarkdown::render("003_kb_fetch_ca.Rmd")
 ```
@@
 # dump
-write_csv(rp_country, "data/rp_data_14_18.csv")
+write_csv(rp_country, "../data/rp_data_14_18.csv")
 ```
@@
 ```{r}
-cr_journals <- readr::read_csv("data/cr_wos.csv") %>%
+cr_journals <- readr::read_csv("../data/cr_wos.csv") %>%
   # merge springer nature brands
   mutate(publisher = ifelse(grepl("Springer", publisher, fixed = FALSE), "Springer Nature", publisher))
-rp_country <- readr::read_csv("data/rp_data_14_18.csv")
-issn_l <- readr::read_tsv("data/20190818.ISSN-to-ISSN-L.txt") %>%
+rp_country <- readr::read_csv("../data/rp_data_14_18.csv")
+issn_l <- readr::read_tsv("../data/20190818.ISSN-to-ISSN-L.txt") %>%
   # manual fix Journal - American Water Works 
@@
     issn_l = `ISSN-L`) %>%
-  write_csv("data/rp_jn_14_18.csv")
+  write_csv("../data/rp_jn_14_18.csv")
 ```
@@
 require(writexl)
-rp_df <- readr::read_csv("data/rp_jn_14_18.csv")
+rp_df <- readr::read_csv("../data/rp_jn_14_18.csv")
 tmp <- map(2014:2018, function(x) {
@@
 names(tmp) <- paste0("pubyear_", 2014:2018)
-write_xlsx(tmp, "data/rp_14_18.xlsx")
+write_xlsx(tmp, "../data/rp_14_18.xlsx")
 ```

### 2019-09-09 analysis/004_kb_fetch_publisher.Rmd
rmd reports render

@@
 title: "Publisher League Global"
-author: "Najko Jahn"
-date: "9/5/2019"
 output: github_document
@@
 require(writexl)
-rmarkdown::render("kb_fetch_publisher_global.Rmd")
+rmarkdown::render("004_kb_fetch_publisher_global.Rmd")
 ```
@@
 jdbcDriver <-
-  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar") # you may need to change the path to your db driver
+  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "../inst/jdbc_driver/ojdbc8.jar") # you may need to change the path to your db driver
 jdbcConnection <-
@@
 # dump
-write_csv(rp_publisher, "data/publisher_league_14_18.csv")
+write_csv(rp_publisher, "../data/publisher_league_14_18.csv")
 ```
@@
 ```{r}
-cr_journals <- readr::read_csv("data/cr_wos.csv") %>%
+cr_journals <- readr::read_csv("../data/cr_wos.csv") %>%
   # merge springer nature brands
   mutate(publisher = ifelse(grepl("Springer", publisher, fixed = FALSE), "Springer Nature", publisher))
-rp_publisher <- readr::read_csv("data/publisher_league_14_18.csv")
-issn_l <- readr::read_tsv("data/20190818.ISSN-to-ISSN-L.txt") %>%
+rp_publisher <- readr::read_csv("../data/publisher_league_14_18.csv")
+issn_l <- readr::read_tsv("../data/20190818.ISSN-to-ISSN-L.txt") %>%
   # manual fix Journal - American Water Works 
@@
     ) %>%
-  write_csv("data/publisher_14_18.csv")
+  write_csv("../data/publisher_14_18.csv")
 ```
@@
 ```{r}
-readr::read_csv("data/publisher_14_18.csv") %>%
-  writexl::write_xlsx("data/publisher_14_18.xlsx")
+readr::read_csv("../data/publisher_14_18.csv") %>%
+  writexl::write_xlsx("../data/publisher_14_18.xlsx")
 ```

### 2019-09-09 analysis/paper.Rmd
rmd reports render

@@
 output: github_document
+bibliography: pubs.bib
 ---
@@
 
-The volume of corresponding author publications is a crucial indicator in the planning of open access funding programs and licensing models (Schimmer, Geschuhn, and Vogler 2015).
+The volume of corresponding author publications is a crucial indicator in the planning of open access funding programs and licensing models [@Schimmer_2015].
 But funders and national library consortia often lack access to such bibliometric data.
@@
 Only a few openly available datasets focusing on the scholarly publishing industry exist.
-For example, Larivière et al.
-(2015), shared an aggregated dataset that contains the market share of publishers in terms of articles indexed in the Web of Science between 1973-2013.
-Using this data, the authors were able to reveal a growing market concentration among a few publishing houses.
-Palzenberger (2015) provided a dataset presenting the number and proportion of scholarly articles per country of affiliation including a breakdown per author role for the period 2004-2014.
+For example, @haustein_2015 shared an aggregated dataset that contains the market share of publishers in terms of articles indexed in the Web of Science between 1973-2013.
+Using this data, the authors were able to reveal a growing market concentration among a few publishing houses [@Larivi_re_2015].
+@Palzenberger_2015 provided a dataset presenting the number and proportion of scholarly articles per country of affiliation including a breakdown per author role for the period 2004-2014.
 Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to open access business model is feasible.
@@
 In these databases, Unpaywall data is often used to identify openly available full-texts at the article-level, while the Directory of Open Access Journals provides information about whether the journal is operating under a fully open access or toll-access business model.
-In its most recent edition, the Leiden Ranking presents open access indicators at the level of universities.
-In addition to university-specific data, the German Open Access Monitor, provides breakdowns per publisher and journal, as well as the inclusion of publications from non-university research institutions based in Germany
+In its most recent edition, the Leiden Ranking presents open access indicators at the level of universities [@leiden].
+In addition to university-specific data, the German Open Access Monitor, provides breakdowns per publisher and journal, as well as the inclusion of publications from non-university research institutions based in Germany [@oa_monitor].
 
 Here, we build an open dataset on the global productivity of corresponding authors.
-Drawing on Palzenberger's previous work, we not only created an updated dataset about the number and proportion of scholarly articles per country of affiliation from corresponding authors.
+Drawing on Palzenberger's (2015) previous work, we not only created an updated dataset about the number and proportion of scholarly articles per country of affiliation from corresponding authors.
 The dataset also allows for a breakdown by journal, including information whether the journal is currently fully open access or not, and publishers.
 We constructed the dataset using the quality assured Web of Science in-house database from the German Competence Center for Bibliometrics.
-We enriched the data with publisher information from Crossref, a large DOI registration agency for scholarly work, and with open access journal information provided by the Biefeld GOLD OA list.
+We enriched the data with publisher information from Crossref, a large DOI registration agency for scholarly work, and with open access journal information provided by the Biefeld GOLD OA list [@issn].
 
@@
 
-Following Palzenberger, data were retrieved from the in-house Web of Science database maintained by the German Competence Center for Bibliometrics. The following data were obtained and aggregated:
+Following @Palzenberger_2015, data were retrieved from the in-house Web of Science database maintained by the German Competence Center for Bibliometrics.
+The following data were obtained and aggregated:
 
@@
 
-Our database retrieval reflected that it has become challenging to count corresponding authors straight, because the Web of Science has begun to keep track of more than one reprint author recently.
+Our database retrieval reflected that it has become challenging to count corresponding authors straight, because the Web of Science has begun to keep track of more than one reprint author recently (see Figure 1).
 While evaluating an efficient retrieval strategy, we found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018.
-Moreover, a percentage of 5% of reprint author was internationally co-located (see Figure 1).
+Moreover, a percentage of 5% of reprint author was internationally co-located (see Figure 2).
 We therefore decided to use whole counting where every Web of Science record representing an original article or review was counted once per country of affiliation.
 
-![](figure/country_wos.png)
+![Figure 1](002_kb_rp_coverage_files/figure-gfm/ca_author_count-1.png)
+
+![Figure 2](002_kb_rp_coverage_files/figure-gfm/ca_country_count-1.png)
 
@@
 ```{r}
-publisher_df <- readr::read_csv("data/publisher_14_18.csv")
+publisher_df <- readr::read_csv("../data/publisher_14_18.csv")
 ```
@@
 ```{r}
-rp_df <- readr::read_csv("data/rp_jn_14_18.csv")
+rp_df <- readr::read_csv("../data/rp_jn_14_18.csv")
 ```
@@
 ```{r, results='asis'}
-my_df <- readr::read_csv("data/publisher_14_18.csv")
+my_df <- readr::read_csv("../data/publisher_14_18.csv")
 tt_all <- my_df %>% 
@@
         legend.justification = "center")
+dir.create("figure")
 ggsave(
@@
 
-![](figure/global_publisher_share.png)
+![Figure 3](figure/global_publisher_share.png)
 
@@
 
-![](figure/worldmap.png)
+```{r}
+my_df <- readr::read_csv("../data/rp_jn_14_18.csv")
+#'  transform to eu
+eu <-
+  c(
+    "AUT",
+    "BEL",
+    "BGR",
+    "CYP",
+    "CZE",
+    "DEU",
+    "DNK",
+    "ESP",
+    "EST",
+    "FIN",
+    "FRA",
+    "GBR",
+    "GRC",
+    "HRV",
+    "HUN",
+    "IRL",
+    "ITA",
+    "LTU",
+    "LUX",
+    "LVA",
+    "MLT",
+    "NLD",
+    "POL",
+    "PRT",
+    "ROU",
+    "SVK",
+    "SVN",
+    "SWE"
+  )
+my_df %>%
+  mutate(eu = case_when(country_code %in% eu ~ "EU28")) %>%
+  mutate(countries = ifelse(eu %in% "EU28", "EU28", country_code)) -> df_to_plot
+#' top 5 publisher closed
+my_df %>%
+  filter(oa_journal == FALSE) %>%
+  group_by(publisher) %>%
+  summarise(n = sum(articles)) %>%
+  arrange(desc(n)) %>%
+  head(7) %>%
+  .$publisher -> top_6
+#' data transformation to prepare plotting of corresponding authors' country of affilation
+df_to_plot %>%
+  mutate(fct_publisher = ifelse(publisher %in% top_6, publisher, "Others")) %>%
+  mutate(fct_publisher = ifelse(is.na(fct_publisher), "Others", fct_publisher)) %>%
+  group_by(fct_publisher, countries) %>%
+  summarize(n = sum(articles)) %>%
+  mutate(prop = n / sum(n)) %>%
+  ungroup() %>%
+  mutate(fct_publisher = factor(fct_publisher, levels = c(top_6[!is.na(top_6)], "Others"))) %>%
+  mutate(countries = fct_reorder(countries, n)) -> plot_df
+#' add country lat lon
+#' eu28 frankfurt main is the center
+eu_df <-
+  data_frame(
+    latitude = 50.6393264,
+    longitude = 8.40556572,
+    name = "EU28",
+    iso3c = "EU28"
+  )
+library(countrycode)
+gc <- readr::read_tsv("../data/geocodes.csv") %>%
+  # add iso code used in kb database
+  mutate(iso3c = countrycode::countrycode(country, "iso2c", "iso3c")) %>%
+  bind_rows(eu_df) %>%
+  filter(!iso3c %in% eu)
+
+gc %>%
+  select(-country) %>%
+  inner_join(plot_df, by = c("iso3c" = "countries")) %>%
+  select(-prop) %>%
+  spread(fct_publisher, n) %>%
+  mutate(all_pubs = rowSums(.[5:11], na.rm = TRUE)) -> plot_wide
+tmp <- plot_wide %>%
+  arrange(desc(all_pubs)) %>%
+  replace(., is.na(.), 0) 
+#' plot
+world <- map_data('world') %>%
+  # without Antartica
+  filter(region != "Antarctica")
+p <- ggplot(world, aes(long, lat)) +
+  geom_map(
+    map = world,
+    aes(map_id = region),
+    color = "white",
+    size = 0.1,
+    fill  = "#b3b3b3a0"
+  ) +
+  coord_map() +
+  scatterpie::geom_scatterpie(
+    aes(
+      x = longitude,
+      y = latitude,
+      r = sqrt(all_pubs / 5000)
+    ),
+    data = tmp,
+    cols = c(levels(plot_df$fct_publisher)),
+    color = NA,
+    alpha = .7
+  ) +
+  geom_text(
+    aes(
+      x = longitude,
+      y = latitude,
+      group = iso3c,
+      label = iso3c
+    ),
+    data = tmp,
+    stat = "identity",
+    size = sqrt(tmp$all_pubs / 50000)
+  ) +
+  scale_fill_manual(
+    "Publisher",
+    values = c(
+      "Elsevier BV" = "#E69F00",
+      "Springer Nature" = "#0072B2",
+      "Wiley" = "#56B4E9",
+      "Informa UK Limited" = "#009E73",
+      "American Chemical Society (ACS)" = "#F0E442",
+      "SAGE Publications" = "#D55E00",
+      "Others" = "#fcfcfc"
+    ),
+    limits = levels(plot_df$fct_publisher)
+  ) +
+  coord_equal() +
+  theme_minimal()  +
+  labs(
+    title = "Who did publish in toll-access journals between 2014 - 2018?",
+    x = "",
+    y = "",
+    caption = "Data Sources:  Web of Science in-house database from the Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata.Closed Access Journals identified using the Bielefeld GOLD OA list.\n
+    Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
+    subtitle = "Circle size represents the number of corresponding authorships of original articles and reviews per country"
+  ) +
+  theme_minimal(base_family = "Roboto") +
+  theme(plot.margin = margin(10, 30, 10, 30)) +
+  theme(panel.grid.minor = element_blank()) +
+  theme(axis.ticks = element_blank()) +
+  theme(panel.grid.major = element_blank()) +
+  theme(panel.border = element_blank()) +
+  theme(axis.ticks = element_blank(),
+        axis.text = element_blank()) +
+  theme(legend.position = "top",
+        legend.justification = "right")
+
+ggsave(
+  "figure/map.png",
+  p,
+  dpi = 450,
+  height = 6,
+  width = 9
+)
+
+```
+
+![Figure 4](figure/map.png)
 

### 2019-09-10 analysis/paper.Rmd
adding @gitti1 improvements

@@
 title: "Data Descriptive: Corresponding author journal publishing activities 2014 - 2018"
-date: "9/5/2019"
 output: github_document
 bibliography: pubs.bib
+csl: chicago.csl
 ---
@@
 
-The volume of corresponding author publications is a crucial indicator in the planning of open access funding programs and licensing models [@Schimmer_2015].
-But funders and national library consortia often lack access to such bibliometric data.
-Particularly, determining the share per publisher, journal and journal business model across countries is challenging.
+Assessing the volume and share of corresponding author publications is crucial for the planning of open access funding programs and licensing models [@Schimmer_2015].
+However, funders and national library consortia often lack access to bibliometric data, which allows for determining the publication output per publisher, journal and journal business model across countries.
 Here, by obtaining country affiliations from the Web of Science at the level of journals, and combining them with normalized publisher and open access status information, we created a comprehensive dataset about the productivity of corresponding authors for the period 2014-2018.
-This dataset is expected to have essential value for the quantitatve understanding of scholarly publishing.
+This dataset is expected to have essential value for the quantitative understanding of scholarly publishing.
 It will not only help to analyze the market share of publishing houses, but will also uncover the current state of the transition of subscription-based publishing models to fully open access.
@@
 @Palzenberger_2015 provided a dataset presenting the number and proportion of scholarly articles per country of affiliation including a breakdown per author role for the period 2004-2014.
-Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to open access business model is feasible.
+Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to open access business model is feasible based on the current total spending on scholarly journals [@Schimmer_2015].
 Since then, analyzing the productivity of corresponding authors in terms of articles published in dedicated journal portfolios has been perceived as a critical component of open access negotiations.
@@
 Drawing on Palzenberger's (2015) previous work, we not only created an updated dataset about the number and proportion of scholarly articles per country of affiliation from corresponding authors.
-The dataset also allows for a breakdown by journal, including information whether the journal is currently fully open access or not, and publishers.
+The dataset also allows for a breakdown by journal, including information whether the journal is currently fully open access or not, and by publishers.
 We constructed the dataset using the quality assured Web of Science in-house database from the German Competence Center for Bibliometrics.
@@
 The data cannot only serve as a critical input for negotiations in the context of transformative agreements, but also in the quantitative study of the academic publishing industry and the open access uptake across countries.
-In the latter case, for instance, national transformative agreements with large publishers will likely have an impact on the open access share of a country.
-To demonstrate its potential significance, we will present two new findings using the dataset: a global map of publishers's toll-access market share, and an analysis of country of affiliation per publisher.
+For instance, national transformative agreements with large publishers will likely have an impact on the open access share of a country in the near future.
+To demonstrate the dataset's potential significance, we will present two new findings using the dataset: a global map of publishers's toll-access market share, and an analysis of country of affiliation by publisher.
 Source code used for compiling the dataset is shared along with the resulting datasets.
@@
 
-Our database retrieval reflected that it has become challenging to count corresponding authors straight, because the Web of Science has begun to keep track of more than one reprint author recently (see Figure 1).
-While evaluating an efficient retrieval strategy, we found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018.
-Moreover, a percentage of 5% of reprint author was internationally co-located (see Figure 2).
-We therefore decided to use whole counting where every Web of Science record representing an original article or review was counted once per country of affiliation.
+Our database retrieval reflected that it has become challenging to count corresponding author affiliations straight where one country receives full credit for the publication [@Huang_2011], because the Web of Science has begun to syetmatically keep track of more than one corresponding author recently (see Figure 1).
+We also found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018.
+Moreover, a  5% of corresponding authors author was internationally co-located (see Figure 2).
+These results support some concerns about how to allocate open access spending based on corresponding authorships [@Gumpenberger_2018].
+Here, we decided to use whole counting where every original article or review was counted once per country of affiliation of the corresponding authors.
 
@@
 
-After obtaining aggregated article counts at the country and journal level form the Web of Science, Crossref’s REST API was queried to retrieve publisher and journal titles.
-To control developments of the publishing market resulting in name changes of publishers or journal titles over time, only the most frequent facet field name was used.
-Drawing on ISSN-L, ISSN variants were obtained to improve the retrieval.
+After obtaining aggregated article counts at the country and journal level from the Web of Science, Crossref’s REST API was queried to retrieve publisher and journal titles.
+To control for developments of the publishing market resulting in name changes of publishers or journal titles over time, only the most frequent field name was used from the API facet counts.
+Drawing on ISSN-L, a jouranl identifier that links between different media versions of a journal, ISSN variants were obtained to improve the retrieval.
 Next, the open access status per journal was added to the dataset.
-This information was retrieved from Bielefeld, which joins open access journal information from various sources including the DOAJ, PubMed Central and the Directory of Open Access Scholarly Resources (ROAD).
+This information was retrieved from a comprehensive list of open access journals indexed in the Web of Science, which joins open access journal information from various sources including the DOAJ, PubMed Central and the Directory of Open Access Scholarly Resources (ROAD) [@issn].
 Again, ISSN-L were used for the matching procedure.
 
+## Data and Code availability
+
 The underlying source code in R and SQL for both testing the retrieval and compiling the datasets are openly shared as executable reports, making our data work more reproducible and transparent.
+The data analytics work, which is organized as research compendium [@Marwick_2017] is accessible at <https://github.com/subugoe/oa2020cadata>
 
-## Data Records
+## The Dataset
 
-Overall, our dataset consists of two separate rectangular data, which are openly available at XXXX. Each data is provided in comma-separated values (csv) files and as Excel spreadsheets.
+Overall, our dataset consists of two separate tables, which are openly available at <https://github.com/subugoe/oa2020cadata> as comma-separated values (csv) files and Excel spreadsheets.
 
-### Global publisher data
+### Global journal and publisher data
 
@@
 
-The first file, named `publisher_14_18`, contains information about the publishing market landscape for the period 2014-2018 in terms of articles and reviews indexed in the Web of Science.
-Table 1 describes data variables in detail.
+The first file, named `publisher_14_18`, contains information about the publishing market landscape for the period 2014-2018 in terms of the aggregated number of articles and reviews  by year and journal represented by the ISSN identifier indexed in the Web of Science.
+Table 1 describes the data variables in detail.
 We were able to retrieve publisher and journal titles from Crossref for `r publisher_df %>% filter(!is.na(publisher)) %>% distinct(issn_l) %>% nrow() / publisher_df %>% distinct(issn_l) %>% nrow() * 100` %
- of journals, representing `r publisher_df %>% filter(!is.na(publisher)) %>% .$articles %>% sum() / sum(publisher_df$articles) * 100` % of the overall article volume.
+ of all journals, representing `r publisher_df %>% filter(!is.na(publisher)) %>% .$articles %>% sum() / sum(publisher_df$articles) * 100` % of the overall article volume.
 `r publisher_df %>% filter(oa_journal == TRUE) %>% distinct(issn_l) %>% nrow()` out of `r publisher_df %>% filter(!is.na(publisher)) %>% distinct(issn_l) %>% nrow()`
@@
 
-|| Variable           | Description                                                                                                                                 | Source                                                                                         |
+| Variable           | Description                                                                                                                                 | Source                                                                                         |
 | :----------------- | :------------------------------------------------------------------------------------------------------------------------------------------ | :--------------------------------------------------------------------------------------------- |
@@
 
-### Corresponding author country-level publisher data
+### Corresponding author data
 
@@
 
-The file named `rp_14_18` provides a breakdown of publisher and journal information per country of affiliation.
+The file named `rp_14_18` provides a breakdown of journal information aggregated by  countries of affiliation.
 Only corresponding authorships, indicated by the role "reprint author" in the Web of Science database, were taken into consideration.
-Table 1 describes data variables in detail.
-Overall, 99 % of records representing original articles and reviews in the period 2014 - 2017 provide affiliation information for reprint authors at the country-level.
+The following table  describes the variables in detail.
+Overall, 99 % of records representing original articles and reviews in the period 2014 - 2017 provide affiliation information for corresponding authors at the country-level.
 
@@
 
-Using the data record `publisher_14_18` the global market share of publishing houses in terms of articles and reviews indexed in the Web of Science was analyzed for the period 2014-2018.
+Based on the dataset `publisher_14_18` the global market share of publishing houses in terms of articles and reviews indexed in the Web of Science can be analyzed for the period 2014-2018.
 Table 1 shows the overall market share for the ten largest publishers.
@@
   arrange(desc(publisher_group))
-knitr::kable(tt_all, caption = "Publisher Market Share 2014 - 2018 in terms of articles and reviews indexed in the Web of Science")
+tt_all %>%
+  mutate(prop = prop * 100) %>%
+  knitr::kable(caption = "Publisher Market Share 2014 - 2018 in terms of articles and reviews indexed in the Web of Science", col.names = c("Publisher", "Articles and Reviews published", "Percentage"))
 ```
@@
 Our data confirms that the global publishing landscape is dominated by a few large publishers.
-In total, ten publishers accounted for 65% of articles published in the period 2018.
+In total, ten publishers accounted foralmost two-thirds of articles published between 2014 - 2018.
 The Big 3, Elsevier, Springer Nature and Wiley, even accounted for a percentage of 46 %.
-However, analysis suggests large differences when differentiating between ftoll and fully open access: The market for fully open access journals is generally less domintated by major publishing houses.
-A notable exeption is Springer Nature with its large fully open access portfolio including BioMed Central or the large mega-journals Scientific Reports and Nature Communications.
+However, further analysis suggests large differences when differentiating between toll and fully open access: The market for fully open access journals is so far less domintated by major publishing houses.
+A notable exeption is Springer Nature with its large fully open access portfolio including BioMed Central and the large mega-journals Scientific Reports and Nature Communications.
 
@@
 
-Using `rp_14_18` the publisher market shares can be furthermore broken down per country of affiliation from corresponding author.
-Figure presents a global map of publishing in toll-access journals.
-Publications from institutions based in the European Union were summarized to "EU28".
-Together, EU 28 member states constituted the largest market for toll-access publishing between 2014-18.
+Using `rp_14_18` the publisher market shares can be further broken down by country affiliation from corresponding authors.
+The following figure presents a global map of publishing in toll-access journals.
+Publications from institutions based in the European Union were summarized as "EU28".
+Together, all EU 28 member states constituted the largest market for toll-access publishing between 2014-18.
 
@@
 
-## Responsible use
+## Responsible use and limitations
 
-Although the Web of Science database is a well-established bibliometric database in the quantitative science studies it comes with certain limitations that needs to considered when interpreting our data.
-First, it consists of a limited set of journals, likely not covering the whole journal portfolio of the investigated publishers.
+Although the Web of Science database is a well-established bibliometric database in the quantitative science studies it comes with certain limitations that need careful consideration when interpreting our data.
+First, the Web of Science is selective in terms of journal inclusion, likely not covering the whole journal portfolio of the investigated publishers.
 Although publisher names were disambiguated, journal title transfers between publishers were not tracked.
@@
 print publication date, was used.
-However, there can be considerable time-lag from online-first publication and issue inclusion.
+However, there can be considerable time-lag between online-first publication and issue inclusion.
 
-As any other bibliometric databases, the data structure can be subject of change.
-In particular, our pretests suggest that the reprint author field used to identify corresponding authorships has undergone a change recently, starting to covering more than one reprint author per article.
+Moreover, as any other bibliometric databases, the data structure can be subject of change.
+In particular, our pretests suggest that the Web of Science author role field used to identify corresponding authorships has undergone a change recently, starting to covering more than one reprint author per article.
 
-The Bielefeld GOLD OA list was used to identify fully open access journals.
-One of its weakness is that it does not contain the date of journal flipping.
-The figures should be therefore interpreted as numbers of articles available in current fully open access journals, instead of published in an fully open access journal at the time of publication.
+The Bielefeld GOLD OA list was used to determine fully open access journals.
+However, it does not contain the date when a subscription-based journal became fully open access.
+The figures should be therefore interpreted numbers of articles available in current fully open access journals, instead of published in an fully open access journal at the time of publication.
+Open access article counts are therefore upper estimates.
 
-Last, but not least, qualitative evaluation is still needed. All code is shared to ensure the technical reproducibility of all steps involved in our data work.
+Last, but not least, further qualitative evaluation of the dataset is still needed.
+Code and data are shared as research compendium to ensure the technical reproducibility of all steps involved in our data analytical work including this data descripitive.
 

### 2019-09-10 analysis/paper.Rmd
fix typo

@@
 | CIEPS                                                                                          |
-                                                                                             |
 
@@
 Last, but not least, further qualitative evaluation of the dataset is still needed.
-Code and data are shared as research compendium to ensure the technical reproducibility of all steps involved in our data analytical work including this data descripitive.
+Code and data are shared as research compendium to ensure the technical reproducibility of all steps involved in our data analytical work including this data descriptive.
 

### 2019-09-13 analysis/004_kb_fetch_publisher.Rmd
adding file name as suggested by  @gitti1

@@
     ) %>%
-  write_csv("../data/publisher_14_18.csv")
+  write_csv("../data/journal_publisher_14_18.csv")
 ```
@@
 ```{r}
-readr::read_csv("../data/publisher_14_18.csv") %>%
-  writexl::write_xlsx("../data/publisher_14_18.xlsx")
+readr::read_csv("../data/journal_publisher_14_18.csv") %>%
+  writexl::write_xlsx("../data/journal_publisher_14_18.xlsx")
 ```

### 2019-09-13 analysis/paper.Rmd
adding file name as suggested by  @gitti1

@@
 Using this data, the authors were able to reveal a growing market concentration among a few publishing houses [@Larivi_re_2015].
-@Palzenberger_2015 provided a dataset presenting the number and proportion of scholarly articles per country of affiliation including a breakdown per author role for the period 2004-2014.
-Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to open access business model is feasible based on the current total spending on scholarly journals [@Schimmer_2015].
+@Palzenberger_2015 provided a dataset presenting the number and proportion of scholarly articles by country of affiliation including a breakdown by author role for the period 2004-2014.
+Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to an open access business model is feasible based on the current total spending on scholarly journals [@Schimmer_2015].
 Since then, analyzing the productivity of corresponding authors in terms of articles published in dedicated journal portfolios has been perceived as a critical component of open access negotiations.
@@
 In its most recent edition, the Leiden Ranking presents open access indicators at the level of universities [@leiden].
-In addition to university-specific data, the German Open Access Monitor, provides breakdowns per publisher and journal, as well as the inclusion of publications from non-university research institutions based in Germany [@oa_monitor].
+In addition to university-specific data, the German Open Access Monitor, provides breakdowns by publisher and journal, as well as the inclusion of publications from non-university research institutions based in Germany [@oa_monitor].
 
@@
 The data cannot only serve as a critical input for negotiations in the context of transformative agreements, but also in the quantitative study of the academic publishing industry and the open access uptake across countries.
-For instance, national transformative agreements with large publishers will likely have an impact on the open access share of a country in the near future.
+For instance, national transformative agreements with large publishers will increase opportunities to publish open access.
+They  will therefore likely have an impact on the open access share of a country with such agreements.
 To demonstrate the dataset's potential significance, we will present two new findings using the dataset: a global map of publishers's toll-access market share, and an analysis of country of affiliation by publisher.
@@
 - Publication years 2014 - 2018 
-- Author Roles
+- Author roles, interpreting the role "reprint author" as corresponding authorship.
 
-Our database retrieval reflected that it has become challenging to count corresponding author affiliations straight where one country receives full credit for the publication [@Huang_2011], because the Web of Science has begun to syetmatically keep track of more than one corresponding author recently (see Figure 1).
+There has been some concerns to allocate open access spending based on corresponding authorships, because it has become challenging to count corresponding author affiliations straight, i.e.
+a counting method where one institution or country receives full credit for a publication [@Huang_2011, @Gumpenberger_2018].
+Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently (see Figure 1).
 We also found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018.
 Moreover, a 5% of corresponding authors author was internationally co-located (see Figure 2).
-These results support some concerns about how to allocate open access spending based on corresponding authorships [@Gumpenberger_2018].
-Here, we decided to use whole counting where every original article or review was counted once per country of affiliation of the corresponding authors.
+We therefore decided to use whole counting where every original article or review was counted once per country of affiliation of the corresponding author(s).
 
@@
 
-After obtaining aggregated article counts at the country and journal level from the Web of Science, Crossref’s REST API was queried to retrieve publisher and journal titles.
+After obtaining aggregated article counts at the level of countries and journals from the Web of Science, Crossref’s REST API was queried to retrieve publisher and journal titles.
 To control for developments of the publishing market resulting in name changes of publishers or journal titles over time, only the most frequent field name was used from the API facet counts.
-Drawing on ISSN-L, a jouranl identifier that links between different media versions of a journal, ISSN variants were obtained to improve the retrieval.
-Next, the open access status per journal was added to the dataset.
+Drawing on ISSN-L, a journal identifier that links between different media versions of a journal, ISSN variants were obtained to improve the retrieval.
+Next, the open access status per journal was identified and added to the dataset.
 This information was retrieved from a comprehensive list of open access journals indexed in the Web of Science, which joins open access journal information from various sources including the DOAJ, PubMed Central and the Directory of Open Access Scholarly Resources (ROAD) [@issn].
@@
 
-Overall, our dataset consists of two separate tables, which are openly available at <https://github.com/subugoe/oa2020cadata> as comma-separated values (csv) files and Excel spreadsheets.
+Overall, our dataset consists of two separate tables, which are openly available at (Link to Zenodo) as comma-separated values (csv) files and Excel spreadsheets.
 
@@
 ```{r}
-publisher_df <- readr::read_csv("../data/publisher_14_18.csv")
+publisher_df <- readr::read_csv("../data/journal_publisher_14_18.csv")
 ```
 
-The first file, named `publisher_14_18`, contains information about the publishing market landscape for the period 2014-2018 in terms of the aggregated number of articles and reviews  by year and journal represented by the ISSN identifier indexed in the Web of Science.
+The first file, named `journal_publisher_14_18`, contains information about the publishing market landscape for the period 2014-2018 in terms of the aggregated number of articles and reviews  by year and journal represented by the ISSN identifier indexed in the Web of Science.
 Table 1 describes the data variables in detail.
@@
 
-Data Schema `publisher_14_18`:
+Data Schema `journal_publisher_14_18`:
 
@@
 
-Based on the dataset `publisher_14_18` the global market share of publishing houses in terms of articles and reviews indexed in the Web of Science can be analyzed for the period 2014-2018.
+Based on the dataset `journal_publisher_14_18` the global market share of publishing houses in terms of articles and reviews indexed in the Web of Science can be analyzed for the period 2014-2018.
 Table 1 shows the overall market share for the ten largest publishers.
@@
 ```{r, results='asis'}
-my_df <- readr::read_csv("../data/publisher_14_18.csv")
+my_df <- readr::read_csv("../data/journal_publisher_14_18.csv")
 tt_all <- my_df %>% 

### 2019-09-13 analysis/paper.Rmd
adding publisher share analysis

@@
        subtitle = "Original Articles and Reviews indexed in the Web of Science\n\n",
-       caption = "\nData Source:  Web of Science / Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata.
-OA journals identified using the Bielefeld Gold OA list.\nFunding: Federal Ministry of Education and Research of Germany (BMBF)") +
+       caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata.
+OA status obtained from the Bielefeld GOLD OA list.\n
+    Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)") +
   #  scale_y_continuous(breaks = seq(0,1, by =0.1), labels = scales::percent_format(accuracy = 5L)) +
@@
 Our data confirms that the global publishing landscape is dominated by a few large publishers.
-In total, ten publishers accounted foralmost two-thirds of articles published between 2014 - 2018.
-The Big 3, Elsevier, Springer Nature and Wiley, even accounted for a percentage of 46 %.
-However, further analysis suggests large differences when differentiating between toll and fully open access: The market for fully open access journals is so far less domintated by major publishing houses.
-A notable exeption is Springer Nature with its large fully open access portfolio including BioMed Central and the large mega-journals Scientific Reports and Nature Communications.
+In total, ten publishers accounted for around two-thirds of articles published between 2014 - 2018.
+The Big 3, Elsevier, Springer Nature and Wiley, accounted for a percentage of 46 %.
+However, further analysis suggests large differences when differentiating between toll and fully open access: The market for fully open access journals seems to be far less dominated by major publishing houses.
+A notable exception is Springer Nature with its large fully open access portfolio including BioMed Central and the large mega-journals Scientific Reports and Nature Communications.
 
@@
     y = "",
-    caption = "Data Sources:  Web of Science in-house database from the Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata.Closed Access Journals identified using the Bielefeld GOLD OA list.\n
+    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata.
+OA status obtained from the Bielefeld GOLD OA list.\n
     Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
@@
 
+The following figure shows the country share per publisher in terms of original articles and reviews published in toll-access journals using corresponding author affiliations.
+The flipped x-axis shows the top 10 countries of affiliations in terms of articles published, while the y-axis depicts the percentage per publisher.
+The figure indicates, which countries were of particularly importance to a publisher.
+Furthermore, it suggests large variations across publishers.
+Take for instance the publisher Royal Society of Chemistry.
+Around every third article published in a toll-access journal was submitted from an author based in China, while the overall share of China-based authors was `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(articles)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
+
+```{r}
+top_10_publisher <- rp_df %>% 
+  filter(oa_journal == FALSE) %>%
+  group_by(publisher) %>%
+  summarise(n = sum(articles)) %>%
+  arrange(desc(n)) %>%
+  head(11) %>%
+  .$publisher 
+top_10_countries <- rp_df %>%
+  filter(oa_journal == FALSE) %>%
+  group_by(country_code) %>%
+  summarise(n = sum(articles)) %>%
+  arrange(desc(n)) %>%
+  head(10) %>%
+  .$country_code
+plot_df <- rp_df %>%
+   filter(oa_journal == FALSE) %>%
+  mutate(fct_publisher = ifelse(publisher %in% top_10_publisher, publisher, "Others")) %>%
+  mutate(fct_publisher = ifelse(is.na(fct_publisher), "Others", fct_publisher)) %>%
+  mutate(fct_countrcode = ifelse(country_code %in% top_10_countries, country_code, "Others")) %>%
+  mutate(fct_countrcode = factor(fct_countrcode, levels = c(top_10_countries, "Others"))) %>%
+  group_by(fct_publisher, fct_countrcode) %>%
+  summarize(n = sum(articles)) %>%
+  mutate(prop = n / sum(n)) %>%
+  ungroup() 
+#' plot
+ggplot(plot_df, aes(forcats::fct_rev(fct_countrcode), prop, fill = prop)) +
+  coord_flip() +
+  geom_bar(stat = "identity") +
+  facet_wrap( ~ fct_relevel(fct_publisher, top_10_publisher), ncol = 3) +
+  scale_fill_viridis_c(
+    "Country share per publisher (in %)",
+    labels = scales::percent_format(accuracy = 5L),
+    option = "C"
+  ) +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 5L))  +
+  labs(
+    title = "Who did publish in toll-access journals from major publishers between 2014 - 2018?",
+    x = "Country of Affiliation (Top 10)",
+    y = "Share of Original Articles and Reviews per Publisher",
+    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata.
+OA status obtained from the Bielefeld GOLD OA list.\n
+    Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
+    subtitle = "Corresponding authors' countries of affiliation per publisher."
+  ) +
+  theme_minimal(base_family = "Roboto", base_size = 16) +
+  theme(plot.margin = margin(30, 10, 30, 10)) +
+  theme(panel.grid.minor.y = element_blank()) +
+  theme(axis.ticks = element_blank()) +
+  theme(panel.grid.major.y = element_blank()) +
+  theme(panel.border = element_blank()) +
+  theme(legend.position = "top",
+        legend.justification = "right",
+        legend.text=element_text(size =  9)) -> p
+#' save as png
+ggsave(
+  "figure/publisher_share.png",
+  p,
+  dpi = 450,
+  height = 12,
+  width = 16
+)
+```
+
+![](figure/publisher_share.png)
+
 ## Responsible use and limitations
 
-Although the Web of Science database is a well-established bibliometric database in the quantitative science studies it comes with certain limitations that need careful consideration when interpreting our data.
+Although the Web of Science database is a well-established bibliometric database in the quantitative science studies it comes with certain limitations that need careful consideration when using and interpreting our data.
 First, the Web of Science is selective in terms of journal inclusion, likely not covering the whole journal portfolio of the investigated publishers.
@@
 print publication date, was used.
-However, there can be considerable time-lag between online-first publication and issue inclusion.
+However, there can be a considerable time-lag between online-first publication and issue inclusion.
 
 Moreover, as any other bibliometric databases, the data structure can be subject of change.
-In particular, our pretests suggest that the Web of Science author role field used to identify corresponding authorships has undergone a change recently, starting to covering more than one reprint author per article.
+In particular, our pretests suggest that the Web of Science author role field used to identify corresponding authorships has undergone a change recently, starting to covering more than one reprint resp.
+corresponding author per article.
 
-The Bielefeld GOLD OA list was used to determine fully open access journals.
+The Bielefeld GOLD OA list [@issn] was used to determine fully open access journals.
 However, it does not contain the date when a subscription-based journal became fully open access.
-The figures should be therefore interpreted numbers of articles available in current fully open access journals, instead of published in an fully open access journal at the time of publication.
+The figures should be therefore interpreted as numbers of articles available in current fully open access journals, instead of published in an fully open access journal at the time of publication.
 Open access article counts are therefore upper estimates.

### 2019-09-13 analysis/paper.Rmd
some tweaks

@@
 ```{r, echo = FALSE, message = FALSE, warning = FALSE}
+library(tidyverse)
+library(janitor)
+library(scatterpie)
+library(knitr)
 knitr::opts_chunk$set(
@@
 )
-library(tidyverse)
 ```
@@
 Based on the dataset `journal_publisher_14_18` the global market share of publishing houses in terms of articles and reviews indexed in the Web of Science can be analyzed for the period 2014-2018.
-Table 1 shows the overall market share for the ten largest publishers.
-Figure 1 presents a breakdown by journal business model, differentiating between toll access and fully open access.
+The following table  shows the overall market share for the ten largest publishers.
+The following figure presents a breakdown by journal business model, differentiating between toll access and fully open access.
 
@@
   summarise(n = sum(articles)) %>%
-  mutate(prop = n /sum(n))  %>%
+  mutate(prop = n / sum(n))  %>%
   mutate(publisher = ifelse(is.na(publisher), "Other", publisher)) %>%
@@
   summarise(n = sum(n), prop = sum(prop)) %>%
-  mutate(publisher_group = fct_reorder(publisher_group, n,)) %>%
+  mutate(publisher_group = fct_reorder(publisher_group, n, )) %>%
   mutate(publisher_group = fct_relevel(publisher_group, "Other")) %>%
@@
   mutate(prop = prop * 100) %>%
-  knitr::kable(caption = "Publisher Market Share 2014 - 2018 in terms of articles and reviews indexed in the Web of Science", col.names = c("Publisher", "Articles and Reviews published", "Percentage"))
+  janitor::adorn_totals() %>%
+  knitr::kable(
+    caption = "Publisher Market Share 2014 - 2018 in terms of articles and reviews indexed in the Web of Science",
+    col.names = c("Publisher", "Articles and Reviews published", "Percentage"),
+    format.args = list(big.mark = ",")
+  )
 ```
@@
   mutate(publisher_group = factor(publisher_group, levels = levels(tt_all$publisher_group)))
-tableau_palette <- rev(c("#ccc94d","#82853b","#86b4a9","#39737c","#ffd94a","#b85a0d","#98d9e4","#3cb7cc","#ffb977","#ff7f0f","#acd98d","#32a251"))
-p_bar <- ggplot(plot_df, aes(x="", y = prop, fill = publisher_group)) +
-  geom_bar(width = 1, stat = "identity", position = position_stack(reverse = FALSE)) +
-  facet_wrap(~label, ncol = 1) +
+tableau_palette <-
+  rev(
+    c(
+      "#ccc94d",
+      "#82853b",
+      "#86b4a9",
+      "#39737c",
+      "#ffd94a",
+      "#b85a0d",
+      "#98d9e4",
+      "#3cb7cc",
+      "#ffb977",
+      "#ff7f0f",
+      "#acd98d",
+      "#32a251"
+    )
+  )
+p_bar <-
+  ggplot(plot_df, aes(x = "", y = prop, fill = publisher_group)) +
+  geom_bar(width = 1,
+           stat = "identity",
+           position = position_stack(reverse = FALSE)) +
+  facet_wrap( ~ label, ncol = 1) +
   coord_flip() +
   geom_text(aes(label = paste0(round(prop, 2) * 100, "")), position = position_stack(vjust = .5)) +
- scale_fill_manual("", values = c("grey80", tableau_palette[1:10]), guide = guide_legend(reverse = TRUE,  ncol= 4)) +
-  scale_y_continuous(breaks = seq(0,1, by =0.1), labels = scales::percent_format(accuracy = 5L)) +
-  labs(title = "Global Publisher Market Shares 2014 - 2018", x = "", y = "",
+  scale_fill_manual(
+    "",
+    values = c("grey80", tableau_palette[1:10]),
+    guide = guide_legend(reverse = TRUE,  ncol = 4)
+  ) +
+  scale_y_continuous(breaks = seq(0, 1, by = 0.1),
+                     labels = scales::percent_format(accuracy = 5L)) +
+  labs(
+    title = "Global Publisher Market Shares 2014 - 2018",
+    x = "",
+    y = "",
     subtitle = "Original Articles and Reviews indexed in the Web of Science\n\n",
@@
 OA status obtained from the Bielefeld GOLD OA list.\n
-    Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)") +
+    Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)"
+  ) +
   #  scale_y_continuous(breaks = seq(0,1, by =0.1), labels = scales::percent_format(accuracy = 5L)) +
@@
   theme(panel.border = element_blank()) +
-  theme(legend.position="bottom",
+  theme(legend.position = "bottom",
         legend.justification = "center")
@@
   height = 9,
-  dpi = 300)
+  dpi = 300
+)
 ```
@@
 
-ggsave(
-  "figure/map.png",
+ggsave("figure/map.png",
        p,
@@
        height = 6,
-  width = 9
-)
-
+       width = 9)
 ```
@@
   geom_bar(stat = "identity") +
-  facet_wrap( ~ fct_relevel(fct_publisher, top_10_publisher), ncol = 3) +
+  facet_wrap(~ fct_relevel(fct_publisher, top_10_publisher), ncol = 3) +
   scale_fill_viridis_c(
@@
   theme(panel.border = element_blank()) +
-  theme(legend.position = "top",
+  theme(
+    legend.position = "top",
     legend.justification = "right",
-        legend.text=element_text(size =  9)) -> p
+    legend.text = element_text(size =  9)
+  ) -> p
 #' save as png
@@
   dpi = 450,
-  height = 12,
-  width = 16
+  height = 13,
+  width = 17
 )

### 2019-10-04 analysis/paper.Rmd
presenting top 10 publsiher with the same color palette

@@
 In its most recent edition, the Leiden Ranking presents open access indicators at the level of universities [@leiden].
-In addition to university-specific data, the German Open Access Monitor, provides breakdowns by publisher and journal, as well as the inclusion of publications from non-university research institutions based in Germany [@oa_monitor].
+In addition to university-specific data, the German Open Access Monitor provides breakdowns by publisher and journal and includes publications from non-university research institutions based in Germany [@oa_monitor].
 
@@
 Drawing on Palzenberger's (2015) previous work, we not only created an updated dataset about the number and proportion of scholarly articles per country of affiliation from corresponding authors.
-The dataset also allows for a breakdown by journal, including information whether the journal is currently fully open access or not, and by publishers.
+The dataset also allows for a breakdown by journal, including information whether it was operated as a subscription or fully open access, and by publishers.
 We constructed the dataset using the quality assured Web of Science in-house database from the German Competence Center for Bibliometrics.
@@
 They  will therefore likely have an impact on the open access share of a country with such agreements.
-To demonstrate the dataset's potential significance, we will present two new findings using the dataset: a global map of publishers's toll-access market share, and an analysis of country of affiliation by publisher.
+To demonstrate the dataset's potential significance, we will present two new findings using the dataset: a global map of publishers's market share, and an analysis of country of affiliation by publisher.
 Source code used for compiling the dataset is shared along with the resulting datasets.
@@
 
-There has been some concerns to allocate open access spending based on corresponding authorships, because it has become challenging to count corresponding author affiliations straight, i.e.
-a counting method where one institution or country receives full credit for a publication [@Huang_2011, @Gumpenberger_2018].
+There have been some concerns to allocate open access spending based on corresponding authorships, because it has become challenging to count corresponding author affiliations straight [@Gumpenberger_2018].
+Under this counting method, only one institution or one country receives full credit for a publication [@Huang_2011].
 Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently (see Figure 1).
@@
   summarise(n = sum(articles)) %>%
-  mutate(oa_journal = ifelse(oa_journal == TRUE, "Open Access", "Toll Access")) %>%
+  mutate(oa_journal = ifelse(oa_journal == TRUE, "Full Open Access", "Subscription")) %>%
   ungroup() %>%
@@
   mutate(publisher_group = factor(publisher_group, levels = levels(tt_all$publisher_group)))
-tableau_palette <-
-  rev(
+london_palette <-
   c(
-      "#ccc94d",
-      "#82853b",
-      "#86b4a9",
-      "#39737c",
-      "#ffd94a",
-      "#b85a0d",
-      "#98d9e4",
-      "#3cb7cc",
-      "#ffb977",
-      "#ff7f0f",
-      "#acd98d",
-      "#32a251"
-    )
+    "#6da7de",
+    "#9e0059",
+    "#dee000",
+    "#d82222",
+    "#5ea15d",
+    "#943fa6",
+    "#63c5b5",
+    "#ff38ba",
+    "#eb861e",
+    "#ee266d"
   )
@@
     "",
-    values = c("grey80", tableau_palette[1:10]),
+    values = c("grey80", rev(london_palette[1:10])),
     guide = guide_legend(reverse = TRUE,  ncol = 4)
@@
 my_df %>%
-  filter(oa_journal == FALSE) %>%
+ # filter(oa_journal == FALSE) %>%
   group_by(publisher) %>%
@@
   arrange(desc(n)) %>%
-  head(7) %>%
+  head(11) %>%
   .$publisher -> top_6
@@
   scale_fill_manual(
-    "Publisher",
-    values = c(
-      "Elsevier BV" = "#E69F00",
-      "Springer Nature" = "#0072B2",
-      "Wiley" = "#56B4E9",
-      "Informa UK Limited" = "#009E73",
-      "American Chemical Society (ACS)" = "#F0E442",
-      "SAGE Publications" = "#D55E00",
-      "Others" = "#fcfcfc"
-    ),
+    "",
+    values = c(london_palette, "#fcfcfc"),
     limits = levels(plot_df$fct_publisher)
@@
   labs(
-    title = "Who did publish in toll-access journals between 2014 - 2018?",
+    title = "Who did publish in scholarly journals between 2014 - 2018?",
     x = "",
     y = "",
-    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata.
-OA status obtained from the Bielefeld GOLD OA list.\n
+    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.
     Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
-    subtitle = "Circle size represents the number of corresponding authorships of original articles and reviews per country"
+    subtitle = "Circle size represents the number of corresponding authorships of original articles and reviews per country\n"
   ) +
@@
         axis.text = element_blank()) +
-  theme(legend.position = "top",
-        legend.justification = "right")
+  theme(legend.position = "bottom",
+        legend.justification = "center") +
+  theme(legend.text = element_text(size = 7.5))
 
@@
        height = 6,
-       width = 9)
+       width = 10)
+ggsave("figure/map.pdf",
+       p,
+       device = cairo_pdf,
+       dpi = 450,
+       height = 6,
+       width = 10)
 ```

### 2019-10-05 analysis/paper.Rmd
some minor fixes

@@
 @Palzenberger_2015 provided a dataset presenting the number and proportion of scholarly articles by country of affiliation including a breakdown by author role for the period 2004-2014.
-Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to an open access business model is feasible based on the current total spending on scholarly journals [@Schimmer_2015].
-Since then, analyzing the productivity of corresponding authors in terms of articles published in dedicated journal portfolios has been perceived as a critical component of open access negotiations.
+Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to an open access business model is is financially feasible in terms of the global amount spent on subscriptions [@Schimmer_2015].
+Since then, analyzing corresponding author publications has become a critical component of open access negotiations between library consortia and publishers.
 
@@
 Major bibliometric databases including the Web of Science, Scopus and Dimension have started to integrate open access status information.
-In these databases, Unpaywall data is often used to identify openly available full-texts at the article-level, while the Directory of Open Access Journals provides information about whether the journal is operating under a fully open access or toll-access business model.
+In these databases, Unpaywall data is often used to identify openly available full-texts at the article-level, while the Directory of Open Access Journals provides information about if a journal is fully open access.
 In its most recent edition, the Leiden Ranking presents open access indicators at the level of universities [@leiden].
@@
 Drawing on Palzenberger's (2015) previous work, we not only created an updated dataset about the number and proportion of scholarly articles per country of affiliation from corresponding authors.
-The dataset also allows for a breakdown by journal, including information whether it was operated as a subscription or fully open access, and by publishers.
-We constructed the dataset using the quality assured Web of Science in-house database from the German Competence Center for Bibliometrics.
-We enriched the data with publisher information from Crossref, a large DOI registration agency for scholarly work, and with open access journal information provided by the Biefeld GOLD OA list [@issn].
+The dataset also allows for a breakdown by journal, including information whether it was operated as a subscription or fully open access journal, and by publishers.
+We constructed the dataset using the quality-assured Web of Science in-house database from the German Competence Center for Bibliometrics.
+We enriched the data with publisher information from Crossref, a large DOI registration agency for scholarly work, and with open access journal information provided by the Biefeld GOLD OA list, a dataset extending open access journal information from the DOAJ with data from other sources [@issn].
 
@@
 For instance, national transformative agreements with large publishers will increase opportunities to publish open access.
-They  will therefore likely have an impact on the open access share of a country with such agreements.
+The existence of such agreements will therefore likely have an impact on the open access share of a country with such agreements.
 To demonstrate the dataset's potential significance, we will present two new findings using the dataset: a global map of publishers's market share, and an analysis of country of affiliation by publisher.
@@
 
-There have been some concerns to allocate open access spending based on corresponding authorships, because it has become challenging to count corresponding author affiliations straight [@Gumpenberger_2018].
+The restriction to the article types original article and reviews reflects that in most open access license negotiations other types of journal content like letters or meeting abstracts are not considered.
+To allocate open access costs to institutions, @Schimmer_2015 proposed to use affiliations from corresponding authors, presuming that corresponding author affiliations can be counted straight.
 Under this counting method, only one institution or one country receives full credit for a publication [@Huang_2011].
+However, there have been some concerns with using corresponding authorship’s data from bibliometric databases for cost analysis recently.
+@Gumpenberger_2018 noted two potential issues: multiple corresponding author affiliations and multiple corresponding authors per article.
 Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently (see Figure 1).
@@
 
-After obtaining aggregated article counts at the level of countries and journals from the Web of Science, Crossref’s REST API was queried to retrieve publisher and journal titles.
+After obtaining aggregated article counts at the level of countries and journals from the Web of Science, Crossref’s REST API was queried using the rcrossref package [@rcrossref] to retrieve publisher and journal titles.
 To control for developments of the publishing market resulting in name changes of publishers or journal titles over time, only the most frequent field name was used from the API facet counts.
@@
 The underlying source code in R and SQL for both testing the retrieval and compiling the datasets are openly shared as executable reports, making our data work more reproducible and transparent.
-The data analytics work, which is organized as research compendium [@Marwick_2017] is accessible at <https://github.com/subugoe/oa2020cadata>
+The data analytics work, which was organized as research compendium [@Marwick_2017] using the holepunch R package [@holepunch], is accessible at <https://github.com/subugoe/oa2020cadata>
 
@@
 The following table  describes the variables in detail.
-Overall, 99 % of records representing original articles and reviews in the period 2014 - 2017 provide affiliation information for corresponding authors at the country-level.
+Overall, 99 % of records representing original articles and reviews in the period 2014 - 2018 had affiliation information for corresponding authors at the country-level.
 
@@
 
-Based on the dataset `journal_publisher_14_18` the global market share of publishing houses in terms of articles and reviews indexed in the Web of Science can be analyzed for the period 2014-2018.
+Based on the dataset `journal_publisher_14_18` the global market share of publishing houses in terms of articles and reviews indexed in the Web of Science can be analyzed for the period 2014-18.
 The following table  shows the overall market share for the ten largest publishers.
-The following figure presents a breakdown by journal business model, differentiating between toll access and fully open access.
+The following figure presents a breakdown by journal business model, differentiating between subscription journals and fully open access.
 
@@
   summarise(n = sum(articles)) %>%
-  mutate(oa_journal = ifelse(oa_journal == TRUE, "Full Open Access", "Subscription")) %>%
+  mutate(oa_journal = ifelse(oa_journal == TRUE, "in Fully Open Access Journals", "in Subscription Journals")) %>%
   ungroup() %>%
@@
 plot_df <- tt_all %>%
-  mutate(oa_journal = "All") %>%
+  mutate(oa_journal = "Overall Article Volume") %>%
   bind_rows(tt_oa) %>%
@@
     subtitle = "Original Articles and Reviews indexed in the Web of Science\n\n",
-    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata.
-OA status obtained from the Bielefeld GOLD OA list.\n
+    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science. 
+    Publisher names were disambiguated using Crossref metadata. OA status obtained from the Bielefeld GOLD OA list.
     Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)"
@@
 The Big 3, Elsevier, Springer Nature and Wiley, accounted for a percentage of 46 %.
-However, further analysis suggests large differences when differentiating between toll and fully open access: The market for fully open access journals seems to be far less dominated by major publishing houses.
+However, further analysis suggests large differences when differentiating between subscription and fully open access journals: The market for fully open access journals seems to be far less dominated by major publishing houses.
 A notable exception is Springer Nature with its large fully open access portfolio including BioMed Central and the large mega-journals Scientific Reports and Nature Communications.
 
-### Global Map of Toll-Access Journal Publishing
+### Global Map of Journal Publishing
 
 Using `rp_14_18` the publisher market shares can be further broken down by country affiliation from corresponding authors.
-The following figure presents a global map of publishing in toll-access journals.
+The following figure presents a global map of scholarly journal publishing represented in the Web of Science.
 Publications from institutions based in the European Union were summarized as "EU28".
-Together, all EU 28 member states constituted the largest market for toll-access publishing between 2014-18.
+Together, all EU 28 member states constituted the largest market for journal publishing between 2014-18.
 
@@
 
-The following figure shows the country share per publisher in terms of original articles and reviews published in toll-access journals using corresponding author affiliations.
+The following figure shows the country share per publisher in terms of original articles and reviews published in subscription journals using corresponding author affiliations.
 The flipped x-axis shows the top 10 countries of affiliations in terms of articles published, while the y-axis depicts the percentage per publisher.
@@
 Take for instance the publisher Royal Society of Chemistry.
-Around every third article published in a toll-access journal was submitted from an author based in China, while the overall share of China-based authors was `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(articles)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
+Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(articles)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
 
@@
   labs(
-    title = "Who did publish in toll-access journals from major publishers between 2014 - 2018?",
+    title = "Who did publish in subscription-based journals from major publishers between 2014 - 2018?",
     x = "Country of Affiliation (Top 10)",
     y = "Share of Original Articles and Reviews per Publisher",
-    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.\nPublisher names were disambiguated using Crossref metadata.
-OA status obtained from the Bielefeld GOLD OA list.\n
+    caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.
+    Publisher names were disambiguated using Crossref metadata. OA status obtained from the Bielefeld GOLD OA list.
     Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
@@
 
-Last, but not least, further qualitative evaluation of the dataset is still needed.
-Code and data are shared as research compendium to ensure the technical reproducibility of all steps involved in our data analytical work including this data descriptive.
+In the end, we present a novel bibliometric dataset representing the journal publishing market with a particular focus on open access.
+Although access to the underlying data infrastructure is restricted, we decided to openly share the aggregated dataset as part of a research compendium to demonstrate the technical reproducibility of this work.
 

### 2019-10-05 analysis/paper.Rmd
some minor fixes

@@
   summarise(n = sum(articles)) %>%
-  mutate(oa_journal = ifelse(oa_journal == TRUE, "in Fully Open Access Journals", "in Subscription Journals")) %>%
+  mutate(oa_journal = ifelse(oa_journal == TRUE, "Fully Open Access", "Subscription")) %>%
   ungroup() %>%
@@
 plot_df <- tt_all %>%
-  mutate(oa_journal = "Overall Article Volume") %>%
+  mutate(oa_journal = "All Journals") %>%
   bind_rows(tt_oa) %>%

### 2019-10-08 analysis/paper.Rmd
some fixes

@@
 ---
-title: "Data Descriptive: Corresponding author journal publishing activities 2014 - 2018"
+title: "Data Descriptor: Corresponding author journal publishing activities 2014 - 2018"
 output: github_document
+author: "Najko Jahn"
 bibliography: pubs.bib
@@
 Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to an open access business model is is financially feasible in terms of the global amount spent on subscriptions [@Schimmer_2015].
-Since then, analyzing corresponding author publications has become a critical component of open access negotiations between library consortia and publishers.
+Since then, analyzing corresponding author publications has become a critical component of open access negotiations between library consortia and publishers [@Geschuhn_2017].
 
@@
 
-The data cannot only serve as a critical input for negotiations in the context of transformative agreements, but also in the quantitative study of the academic publishing industry and the open access uptake across countries.
+The data cannot only serve as a critical input for negotiations in the context of transformative agreements, but also in the quantitative study of the scholarly publishing industry and the open access uptake across countries.
 For instance, national transformative agreements with large publishers will increase opportunities to publish open access.
@@
 
-## Data and Methods
+## Methods
 
@@
 
-The restriction to the article types original article and reviews reflects that in most open access license negotiations other types of journal content like letters or meeting abstracts are not considered.
+Like @Palzenberger_2015The, above-listed selection reflect data needs in the context of open access negotiations between library consortia and publishers.
+The restriction to the article types original article and reviews is in accordance with most open access contracts where other types of journal content like letters or meeting abstracts are not considered (see for instance @Sander_2019).
+
 To allocate open access costs to institutions, @Schimmer_2015 proposed to use affiliations from corresponding authors, presuming that corresponding author affiliations can be counted straight.
@@
 @Gumpenberger_2018 noted two potential issues: multiple corresponding author affiliations and multiple corresponding authors per article.
-Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently (see Figure 1).
-We also found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018.
+
+Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently: We found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018 (see Figure 1).
 Moreover, a 5% of corresponding authors author was internationally co-located (see Figure 2).
@@
 The underlying source code in R and SQL for both testing the retrieval and compiling the datasets are openly shared as executable reports, making our data work more reproducible and transparent.
-The data analytics work, which was organized as research compendium [@Marwick_2017] using the holepunch R package [@holepunch], is accessible at <https://github.com/subugoe/oa2020cadata>
+The data analytics work, which was organized as research compendium [@Marwick_2017] using the holepunch R package [@holepunch], is accessible at <https://github.com/subugoe/oa2020cadata>.
+
+More specifically, the following database retrieval steps were tracked using R Markdown, a reproducible authoring framework supporting various programming languages including R and SQL, and are accessible in the `analysis` folder of the research compendium:
+
+1. Initial exploration about how corresponding authorships are represented in the WoS-KB ([`001_kb_rp_pretest.Rmd`](analysis/001_kb_rp_pretest.Rmd)) 
+2. Coverage analysis of reprint authors in the WoS-KB for the period 2014 - 2018 ([`002_kb_rp_coverage.Rmd`](analysis/002_kb_rp_coverage.Rmd)). 
+3. Steps to obtain and compile the dataset about corresponding author country affiliations 2014 - 2018([`003_kb_fetch_ca.Rmd`](analysis/003_kb_fetch_ca.Rmd)) 
+4. Steps to obtain and compile the dataset about the global publisher output 2014 - 2018 ([`004_kb_fetch_publisher.Rmd`](analysis/004_kb_fetch_publisher.Rmd)).
+
+The research compendium contains the rendered reports as well.
+Furthermore, this source code repository includes comprehensive reproducibility guidance for users with access to the in-house data infrastructure provided by the Competence Centre for Bibliometrics for German Science.
 
@@
 
-Overall, our dataset consists of two separate tables, which are openly available at (Link to Zenodo) as comma-separated values (csv) files and Excel spreadsheets.
+Overall, our dataset consists of the following two separate tables:
+
+-- `journal_publisher_14_18`: Global journal and publisher data 2014-18
+-- `rp_jn_14_18`: Corresponding author country affiliations per publisher and journal 2014-18
+
+They are provided as comma-separated values (csv) files and Excel spreadsheets.
+
+The data are openly available in the `data/` of the research compendium hosted on GitHub <https://github.com/subugoe/oa2020cadata>.
 
@@
 
-The file named `rp_14_18` provides a breakdown of journal information aggregated by  countries of affiliation.
+The file named `rp_jn_14_18` provides a breakdown of journal information aggregated by  countries of affiliation.
 Only corresponding authorships, indicated by the role "reprint author" in the Web of Science database, were taken into consideration.
@@
 In total, ten publishers accounted for around two-thirds of articles published between 2014 - 2018.
-The Big 3, Elsevier, Springer Nature and Wiley, accounted for a percentage of 46 %.
-However, further analysis suggests large differences when differentiating between subscription and fully open access journals: The market for fully open access journals seems to be far less dominated by major publishing houses.
+The Big 3, Elsevier, Springer Nature and Wiley, secured a market share of 46 %.
+However, further analysis suggests large differences between subscription and fully open access journals: The market for fully open access journals seems to be less dominated by major publishing houses.
 A notable exception is Springer Nature with its large fully open access portfolio including BioMed Central and the large mega-journals Scientific Reports and Nature Communications.
@@
 
-Although the Web of Science database is a well-established bibliometric database in the quantitative science studies it comes with certain limitations that need careful consideration when using and interpreting our data.
+Although the Web of Science database is a well-established bibliometric database it has certain limitations that need careful consideration when using and interpreting our data.
 First, the Web of Science is selective in terms of journal inclusion, likely not covering the whole journal portfolio of the investigated publishers.

### 2019-10-10 analysis/wos_cr_match.Rmd
redo crossref journal match

---
title: "WoS Crossref Journal Match"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

## Retrieve WoS-KB journal infos

Connect to database

```{r}
# deal with rJava memory allocation 
# https://stackoverflow.com/questions/34624002/r-error-java-lang-outofmemoryerror-java-heap-space
options(java.parameters = "-Xmx1024m")
require(tidyverse)
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "../inst/jdbc_driver/ojdbc8.jar")
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    Sys.getenv("kb_user"),
    Sys.getenv("kb_pwd")
  ) 
```

Search criteria

- Database: wos_b_2019
- Document Types: Articles and Reviews
- Database Collections: `WOS.SCI`, `WOS.SSCI`, `WOS.AHCI`
- Publication Period 2014 - 2018

#### Query

```{sql connection=jdbcConnection, output.var="wos_jn"}
select
        wos_b_2019.items.pubyear,
        wos_b_2019.issues.issn,
        count(distinct(ut_eid)) as pubs                         
    from
        wos_b_2019.items                                            
    inner join
        wos_b_2019.databasecollection                                                                                          
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items                                  
    inner join
        wos_b_2019.issues                                                        
            on wos_b_2019.issues.pk_issues = wos_b_2019.items.fk_issues                             
    where
        wos_b_2019.databasecollection.edition_value in (
            'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'                                                   
        )                                                                              
        and wos_b_2019.items.doctype in (
            'Article', 'Review'                                                   
        )                                                                               
        and wos_b_2019.items.pubyear in (
            2014, 2015, 2016, 2017, 2018                                                   
        )                                                  
    group by
        wos_b_2019.items.pubyear,
        wos_b_2019.issues.issn
```

Backup

```{r}
readr::write_csv(wos_jn, "../data/wos_jns.csv")
```

#### Populate with ISSN info including ISSN-L

```{r}
# load issn l list
wos_jn <- readr::read_csv("../data/wos_jns.csv")
issn_l <- readr::read_tsv("../data/20190818.ISSN-to-ISSN-L.txt")
issn_wos_match <- wos_jn %>%
  distinct(ISSN) %>%
  left_join(issn_l, by = c("ISSN")) 
issn_variants <-
  issn_wos_match %>% 
  inner_join(issn_l, by = c("ISSN-L")) %>%
  distinct(`ISSN-L`, ISSN = ISSN.y)
issn_variants
```

#### Crossref Match

Per every ISSN-L, obtain all ISSN variants and query Crossref for most frequent publisher and journal title

Generate queries

```{r}
issns_list <-
  purrr::map(unique(issn_variants$`ISSN-L`), function(x) {
    issn_l <- x
    issns <- issn_variants %>%
      filter(`ISSN-L` %in% issn_l) %>%
      .$ISSN
    names(issns) <- rep("issn", length(issns))
    issns
  })
```

Define and call function

```{r}
#' search crossref
require(rcrossref)
jn_facets <- purrr::map(issns_list[66], .f = purrr::safely(function(x) {
  tt <- rcrossref::cr_works(
    filter = c(
      x,
      from_pub_date = "2014-01-01",
      until_pub_date = "2018-12-31",
      type = "journal-article"
    ),
    # being explicit about facets improves API performance!
    facet = "container-title:*,publisher-name:*",
    # less api traffic
    select = "DOI"
  )
  #' Parse the relevant information
  #' - `journal_title` - Crossref journal title (in case of journal name change, we use the most frequent name)
  #' - `publisher` - Crossref publisher (in case of publisher name change, we use the most frequent name)
  #'
  #' To Do: switch to current potential
  if (!is.null(tt)) {
    tibble::tibble(
      issn = list(x),
      journal_title = tt$facets$`container-title`$.id[1],
      publisher = tt$facets$publisher$.id[1]
    )
  } else {
    NULL
  }
}))
```

Generate queries

```{r}
issns_list <-
  purrr::map(unique(issn_variants$`ISSN-L`), function(x) {
    issn_l <- x
    issns <- issn_variants %>%
      filter(`ISSN-L` %in% issn_l) %>%
      .$ISSN
    names(issns) <- rep("issn", length(issns))
    issns
  })
```

#' Dump:
jn_facets_df <- purrr::map_df(jn_facets, "result")
jsonlite::stream_out(jn_facets_df, file("../data/jn_facets_df.json"))

### 2019-10-11 analysis/wos_cr_match.Rmd
crossref journal and publisher name match

@@
 require(rcrossref)
-jn_facets <- purrr::map(issns_list[66], .f = purrr::safely(function(x) {
+jn_facets <- plyr::llply(issns_list, purrr::safely(function(x) {
   tt <- rcrossref::cr_works(
@@
   }
-}))
+}), .progress = "text")
 ```
 
-Generate queries
+```{r}
+#' Dump:
+jn_facets_df <- purrr::map_df(jn_facets, "result")
+cr_jn <- jn_facets_df %>% 
+  unnest(issn)
+readr::write_csv(cr_jn, "../data/cr_wos.csv")
+```
+
+#### ADD ISSN-L
 
 ```{r}
-issns_list <-
-  purrr::map(unique(issn_variants$`ISSN-L`), function(x) {
-    issn_l <- x
-    issns <- issn_variants %>%
-      filter(`ISSN-L` %in% issn_l) %>%
-      .$ISSN
-    names(issns) <- rep("issn", length(issns))
-    issns
-  })
+cr_journals <- readr::read_csv("../data/cr_wos.csv") %>%
+  # merge springer nature brands
+  mutate(publisher = ifelse(grepl("Springer", publisher, fixed = FALSE), "Springer Nature", publisher))
+rp_country <- readr::read_csv("../data/rp_data_14_18.csv")
+issn_l <- readr::read_tsv("../data/20190818.ISSN-to-ISSN-L.txt") %>%
+  # manual fix Journal - American Water Works 
+  add_row(ISSN = "2164-4535", `ISSN-L` = "0003-150X")
+rp_df <- rp_country %>%
+  left_join(issn_l, by = "ISSN") %>%
+  left_join(cr_journals, by = c(ISSN = "issn")) %>%
+  distinct()
 ```
 
-#' Dump:
-jn_facets_df <- purrr::map_df(jn_facets, "result")
-jsonlite::stream_out(jn_facets_df, file("../data/jn_facets_df.json"))
+#### Add OA info
+
+```{r}
+u <- "https://pub.uni-bielefeld.de/download/2934907/2934908/ISSN_Gold-OA_3.0.csv"
+bie_oa <- readr::read_csv(u) %>%
+  # remove missing entries with missing ISSN_L
+  filter(!is.na(ISSN_L))
+# add info to rp_df
+rp_df_oa <- rp_df %>%
+  mutate(oa_journal = `ISSN-L` %in% bie_oa$`ISSN_L`)
+```

### 2019-10-11 analysis/003_kb_fetch_ca.Rmd
re-render rp dataset including updated crossref  journal info

@@
   left_join(issn_l, by = "ISSN") %>%
-  left_join(cr_journals, by = c(`ISSN-L`= "issn_l")) %>%
-  select(-issn) %>%
+  left_join(cr_journals, by = c(ISSN = "issn")) %>%
   distinct()
@@
 names(tmp) <- paste0("pubyear_", 2014:2018)
-write_xlsx(tmp, "../data/rp_14_18.xlsx")
+write_xlsx(tmp, "../data/rp_jn_14_18.xlsx")
 ```

### 2019-10-11 analysis/004_kb_fetch_publisher.Rmd
update to new cr match + some fixes from ntaubert

@@
   mutate(publisher = ifelse(grepl("Springer", publisher, fixed = FALSE), "Springer Nature", publisher))
-rp_publisher <- readr::read_csv("../data/publisher_league_14_18.csv")
 issn_l <- readr::read_tsv("../data/20190818.ISSN-to-ISSN-L.txt") %>%

### 2019-10-11 analysis/paper.Rmd
update to new cr match + some fixes from ntaubert

@@
 However, funders and national library consortia often lack access to bibliometric data, which allows for determining the publication output per publisher, journal and journal business model across countries.
-Here, by obtaining country affiliations from the Web of Science at the level of journals, and combining them with normalized publisher and open access status information, we created a comprehensive dataset about the productivity of corresponding authors for the period 2014-2018.
+Here, by obtaining country affiliations from the Web of Science at the level of journals, and combining them with normalized publisher and open access status information, we created a dataset about the productivity of corresponding authors for the period 2014-2018.
 This dataset is expected to have essential value for the quantitative understanding of scholarly publishing.
@@
 @Palzenberger_2015 provided a dataset presenting the number and proportion of scholarly articles by country of affiliation including a breakdown by author role for the period 2004-2014.
-Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to an open access business model is is financially feasible in terms of the global amount spent on subscriptions [@Schimmer_2015].
+Published in 2015, this dataset, which was obtained from the Web of Science in-house database from the German Competence Center for Bibliometrics, formed the empirical basis for demonstrating that a large-scale transition from a subscription-based to an open access business model would be financially feasible in terms of the global amount spent on subscriptions [@Schimmer_2015].
 Since then, analyzing corresponding author publications has become a critical component of open access negotiations between library consortia and publishers [@Geschuhn_2017].
@@
 We constructed the dataset using the quality-assured Web of Science in-house database from the German Competence Center for Bibliometrics.
-We enriched the data with publisher information from Crossref, a large DOI registration agency for scholarly work, and with open access journal information provided by the Biefeld GOLD OA list, a dataset extending open access journal information from the DOAJ with data from other sources [@issn].
+We enriched the data with publisher information from Crossref, a large DOI registration agency for scholarly work, and with open access journal information provided by the ISSN-GOLD-OA list, a dataset that merges open access journal information from the DOAJ, OpenAPC, PubMed and ROAD [@issn].
 
-The data cannot only serve as a critical input for negotiations in the context of transformative agreements, but also in the quantitative study of the scholarly publishing industry and the open access uptake across countries.
+The data does not only provide background information for negotiations in the context of transformative agreements, but can also be used in the quantitative study of the scholarly publishing industry and the open access uptake across countries.
 For instance, national transformative agreements with large publishers will increase opportunities to publish open access.
 The existence of such agreements will therefore likely have an impact on the open access share of a country with such agreements.
-To demonstrate the dataset's potential significance, we will present two new findings using the dataset: a global map of publishers's market share, and an analysis of country of affiliation by publisher.
+To demonstrate the dataset's significance, we will present two new findings using the dataset: a global map of publishers's market share, and an analysis of country of affiliation by publisher.
 Source code used for compiling the dataset is shared along with the resulting datasets.
@@
 
-Like @Palzenberger_2015The, above-listed selection reflect data needs in the context of open access negotiations between library consortia and publishers.
+Like @Palzenberger_2015, the above-listed selection reflect data needs in the context of open access negotiations between library consortia and publishers.
 The restriction to the article types original article and reviews is in accordance with most open access contracts where other types of journal content like letters or meeting abstracts are not considered (see for instance @Sander_2019).
@@
 Drawing on ISSN-L, a journal identifier that links between different media versions of a journal, ISSN variants were obtained to improve the retrieval.
-Next, the open access status per journal was identified and added to the dataset.
-This information was retrieved from a comprehensive list of open access journals indexed in the Web of Science, which joins open access journal information from various sources including the DOAJ, PubMed Central and the Directory of Open Access Scholarly Resources (ROAD) [@issn].
+Next, the open access status per journal was identified and added to the dataset using the ISSN-Gold -OA 3.0 list [@issn].
 Again, ISSN-L were used for the matching procedure.
@@
 
--- `journal_publisher_14_18`: Global journal and publisher data 2014-18
--- `rp_jn_14_18`: Corresponding author country affiliations per publisher and journal 2014-18
+- `journal_publisher_14_18`: Global journal and publisher data 2014-18
+- `rp_jn_14_18.csv`: Corresponding author country affiliations per publisher and journal 2014-18
 
@@
 | `oa_journal`       | Is the journal publishing all articles open access without delay (full open access)?
-| Bielefeld GOLD OA List V3                                                                      |
+| ISSN-GOLD_OA List 3.0                                                                     |
 | `issn_l`           | Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.
@@
 | `oa_journal`       | Is the journal publishing all articles open access without delay (full open access)?
-| Bielefeld GOLD OA List V3                                                                                                                                  |
+| ISSN-GOLD_OA List 3.0                                                                                                                              |
 | `issn_l`           | Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.
@@
 Although the Web of Science database is a well-established bibliometric database it has certain limitations that need careful consideration when using and interpreting our data.
-First, the Web of Science is selective in terms of journal inclusion, likely not covering the whole journal portfolio of the investigated publishers.
+First, the Web of Science is selective in terms of journal inclusion and is not covering the whole journal portfolio of the investigated publishers.
 Although publisher names were disambiguated, journal title transfers between publishers were not tracked.
@@
 
-The Bielefeld GOLD OA list [@issn] was used to determine fully open access journals.
-However, it does not contain the date when a subscription-based journal became fully open access.
-The figures should be therefore interpreted as numbers of articles available in current fully open access journals, instead of published in an fully open access journal at the time of publication.
+The ISSN GOLD-OA list [@issn] was used to determine fully open access journals.
+However, it does not contain information when a subscription-based journal became fully open access.
+The figures should be therefore interpreted as numbers of articles available in current fully open access journals, instead of published in a fully open access journal at the time of publication.
 Open access article counts are therefore upper estimates.

### 2019-10-11 analysis/paper.Rmd
minor

@@
 ---
-title: "Data Descriptor: Corresponding author journal publishing activities 2014 - 2018"
+title: "Data Descriptor: Corresponding author journal publishing activities 2014-2018"
 output: github_document
-author: "Najko Jahn"
 bibliography: pubs.bib

### 2019-10-19 analysis/003_kb_fetch_ca.Rmd
remove journal-level data

@@
 
-- Publisher and journal titles from Crossref
-- Journal open access status provided by the Bielefeld GOLD OA list
+- Publisher information from Crossref
+- Journal open access status provided by the ISSN GOLD OA list
 
@@
 
-### 4. Data coding and dump
+### 4. Data coding and dump journal-level
 
@@
 |CIEPS                                                                                                                                                        |
+
+### 5. Publisher-level data
+
+```{r}
+rp_jn <- readr::read_csv("../data/rp_jn_14_18.csv") %>%
+  group_by(country_code, publication_year, publisher, oa_journal) %>%
+  summarise(n_publications = sum(articles, na.rm = TRUE),
+            n_journals = n_distinct(issn_l))
+# export to csv and excel
+readr::write_csv(rp_jn, "../data/rp_publisher_14_18.csv")
+writexl::write_xlsx(rp_jn, "../data/rp_publisher_14_18.xlsx")
+
+```
+
+Data Schema:
+
+|Variable           |Description                                                                                                                                 |Source                                                                        |
+|:------------------|:-------------------------------------------------------------------------------------------------------------------------------------------|:-----------------------------------------------------------------------------|
+|`country_code`     |Country of affiliation corresponding author, represented as ISO 3 code                                                                      |KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode` |
+|`publication_year` |Year of publication, obtained from KB Web of Science                                                                                        |KB Web of Science: `wos_b_2019.items.pubyear`                                 |
+|`publisher`        |Most frequently used publisher name in terms of articles published between 2014 - 2018.
+If missing, the journal was not indexed in Crossref |Crossref                                                                      |
+|`oa_journal`       |Is the journal publishing all articles open access without delay (full open access)?
+|ISSN GOLD OA List V3                                                          |
+|`n_publications`   |Number of original articles and reviews published                                                                                           |Aggregate data                                                                |
+|`n_journals`       |Number of distinct journals with corresponding publication output                                                                           |Aggregate data                                                                |

### 2019-10-19 analysis/paper.Rmd
update data descriptor

@@
 library(knitr)
+library(scales)
+library(countrycode)
 knitr::opts_chunk$set(
@@
 2. Coverage analysis of reprint authors in the WoS-KB for the period 2014 - 2018 ([`002_kb_rp_coverage.Rmd`](analysis/002_kb_rp_coverage.Rmd)). 
-3. Steps to obtain and compile the dataset about corresponding author country affiliations 2014 - 2018([`003_kb_fetch_ca.Rmd`](analysis/003_kb_fetch_ca.Rmd)) 
+3. Steps to obtain and compile the dataset about corresponding author country affiliations 2014 - 2018 ([`003_kb_fetch_ca.Rmd`](analysis/003_kb_fetch_ca.Rmd)) 
 4. Steps to obtain and compile the dataset about the global publisher output 2014 - 2018 ([`004_kb_fetch_publisher.Rmd`](analysis/004_kb_fetch_publisher.Rmd)).
@@
 - `journal_publisher_14_18`: Global journal and publisher data 2014-18
-- `rp_jn_14_18.csv`: Corresponding author country affiliations per publisher and journal 2014-18
+- `rp_publisher_14_18.csv`: Corresponding author country affiliations per publisher and journal 2014-18
 
@@
 ```{r}
-rp_df <- readr::read_csv("../data/rp_jn_14_18.csv")
+rp_df <- readr::read_csv("../data/rp_publisher_14_18.csv")
 ```
 
-The file named `rp_jn_14_18` provides a breakdown of journal information aggregated by  countries of affiliation.
+The dataset named `rp_publisher_14_18` provides a breakdown of journal information aggregated by  countries of affiliation.
 Only corresponding authorships, indicated by the role "reprint author" in the Web of Science database, were taken into consideration.
@@
 
-Data
-Schema:
+Data Schema:
 
-| Variable           | Description                                                                                                                                                                      | Source                                                                                                                                                     |
-| :----------------- | :------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- | :--------------------------------------------------------------------------------------------------------------------------------------------------------- |
-| `issn_wos`         | ISSN, a standardized journal id.
-| KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                |
-| `country_code`     | Country of affiliation corresponding author, represented as ISO 3 code                                                                                                           | KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode`                                                                              |
-| `publication_year` | Year of publication, obtained from KB Web of Science                                                                                                                             | KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                              |
-| `articles`         | Number of original articles and reviews published.
-Whole counting where internationally co-located corresponding authorships were assigned to each contributing country equally.
-| KB Web of Science: Grouped counts over `wos_b_2019.issues.issn`, `wos_b_2019.d_items_authors_institutions.inst_countrycode` and `wos_b_2019.items.pubyear` |
-| `journal_title`    | Most frequently used journal title in terms of articles published between 2014 - 2018.
-If missing, the journal was not indexed in Crossref                                       | Crossref                                                                                                                                                   |
-| `publisher`        | Most frequently used publisher name in terms of articles published between 2014 - 2018.
-If missing, the journal was not indexed in Crossref                                      | Crossref                                                                                                                                                   |
-| `oa_journal`       | Is the journal publishing all articles open access without delay (full open access)?
-| ISSN-GOLD_OA List 3.0                                                                                                                              |
-| `issn_l`           | Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.
-| CIEPS                                                                                                                                                      |
+|Variable           |Description                                                                                                                                 |Source                                                                        |
+|:------------------|:-------------------------------------------------------------------------------------------------------------------------------------------|:-----------------------------------------------------------------------------|
+|`country_code`     |Country of affiliation corresponding author, represented as ISO 3 code                                                                      |KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode` |
+|`publication_year` |Year of publication, obtained from KB Web of Science                                                                                        |KB Web of Science: `wos_b_2019.items.pubyear`                                 |
+|`publisher`        |Most frequently used publisher name in terms of articles published between 2014 - 2018.
+If missing, the journal was not indexed in Crossref |Crossref                                                                      |
+|`oa_journal`       |Is the journal publishing all articles open access without delay (full open access)?
+|ISSN GOLD OA List V3                                                          |
+|`n_publications`   |Number of original articles and reviews published                                                                                           |Aggregate data                                                                |
+|`n_journals`       |Number of distinct journals with corresponding publication output                                                                           |Aggregate data
 
@@
 
-Using `rp_14_18` the publisher market shares can be further broken down by country affiliation from corresponding authors.
+Using `rp_publisher_14_18.csv` the publisher market shares can be further broken down by country affiliation from corresponding authors.
 The following figure presents a global map of scholarly journal publishing represented in the Web of Science.
@@
 ```{r}
-my_df <- readr::read_csv("../data/rp_jn_14_18.csv")
 #'  transform to eu
@@
   )
-my_df %>%
+rp_df %>%
   mutate(eu = case_when(country_code %in% eu ~ "EU28")) %>%
@@
 #' top 5 publisher closed
-my_df %>%
+rp_df %>%
  # filter(oa_journal == FALSE) %>%
   group_by(publisher) %>%
-  summarise(n = sum(articles)) %>%
+  summarise(n = sum(n_publications)) %>%
   arrange(desc(n)) %>%
@@
   group_by(fct_publisher, countries) %>%
-  summarize(n = sum(articles)) %>%
+  summarize(n = sum(n_publications)) %>%
   mutate(prop = n / sum(n)) %>%
@@
 eu_df <-
-  data_frame(
+  tibble(
     latitude = 50.6393264,
@@
 Take for instance the publisher Royal Society of Chemistry.
-Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(articles)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
+Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(n_publications)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
 
@@
   group_by(publisher) %>%
-  summarise(n = sum(articles)) %>%
+  summarise(n = sum(n_publications)) %>%
   arrange(desc(n)) %>%
@@
   group_by(country_code) %>%
-  summarise(n = sum(articles)) %>%
+  summarise(n = sum(n_publications)) %>%
   arrange(desc(n)) %>%
@@
   group_by(fct_publisher, fct_countrcode) %>%
-  summarize(n = sum(articles)) %>%
+  summarize(n = sum(n_publications)) %>%
   mutate(prop = n / sum(n)) %>%
@@
 
+This work uses Web of Science data by Clarivate Analytics provided by the German Competence Center for Bibliometrics for research purposes.
+
+This work is supported by the Federal Ministry of Education and Research of Germany (BMBF) in the framework Quantitative Research on the Science Sector (Project: "OAUNI Entwicklung und Einflussfaktoren des Open-Access-Publizierens an Universitäten in Deutschland", Förderkennzeichen: 01PU17023A).
+
 ## References

### 2019-10-21 analysis/002_kb_rp_coverage.Rmd
revised paper.Rmd

@@
 
-[An earlier exploration](kb_rp_exploration.md) studying how affiliations from reprint authors are represented in the WOS-KB revealed that only one affiliation is tagged with `RP in the table `wos_b_2019.items_author_institutions` per author.
+[An earlier exploration](kb_rp_exploration.md)<!--link ist tot!--> studying how affiliations from reprint authors are represented in the WOS-KB revealed that only one affiliation is tagged with `RP in the table `wos_b_2019.items_author_institutions` per author.
 To obtain all affiliations per reprint author, the following retrieval strategy can be used:

### 2019-10-21 analysis/paper.Rmd
revised paper.Rmd

--- 2019-10-11 analysis/paper.Rmd
@@
 Assessing the volume and share of corresponding author publications is crucial for the planning of open access funding programs and licensing models [@Schimmer_2015].
-However, funders and national library consortia often lack access to bibliometric data, which allows for determining the publication output per publisher, journal and journal business model across countries.
+However, funders and national library consortia often lack access to bibliometric data, which allow for determining the publication output per publisher, journal and journal business model across countries.
 Here, by obtaining country affiliations from the Web of Science at the level of journals, and combining them with normalized publisher and open access status information, we created a dataset about the productivity of corresponding authors for the period 2014-2018.
@@
 The existence of such agreements will therefore likely have an impact on the open access share of a country with such agreements.
-To demonstrate the dataset's significance, we will present two new findings using the dataset: a global map of publishers's market share, and an analysis of country of affiliation by publisher.
-Source code used for compiling the dataset is shared along with the resulting datasets.
+To demonstrate the dataset's significance, we will present two new findings using the dataset: a global map of publishers' market shares, and an analysis of country of affiliation by publisher.
+The source code used for compiling the dataset is shared along with the resulting datasets.
 
@@
 
-- Web of Science collections SCI, SSCI and AHCI
+- Web of Science collections SCI, SSCI and AHCI <!-- why only these collections?-->
 - Article types Original Articles and Reviews
-- Country of affiliations
+- Countries of affiliation
 - ISSN
@@
 
-Like @Palzenberger_2015, the above-listed selection reflect data needs in the context of open access negotiations between library consortia and publishers.
-The restriction to the article types original article and reviews is in accordance with most open access contracts where other types of journal content like letters or meeting abstracts are not considered (see for instance @Sander_2019).
+Like @Palzenberger_2015, the above-listed selection reflects data needs in the context of open access negotiations between library consortia and publishers.
+The restriction to the article types original article and review is in accordance with most open access contracts where other types of journal content like letters or meeting abstracts are not considered (see for instance @Sander_2019).
 
@@
 
-Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently: We found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018 (see Figure 1).
+Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently: <!-- maybe stress that before 2016 only one unique corresponding author was allowed?--> We found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018 (see Figure 1).
 Moreover, a 5% of corresponding authors author was internationally co-located (see Figure 2).
-We therefore decided to use whole counting where every original article or review was counted once per country of affiliation of the corresponding author(s).
+We therefore<!--delete therefore?--> decided to use whole counting where every original article or review was counted once per country of affiliation of the corresponding author(s).
 
@@
 
-![Figure 2](002_kb_rp_coverage_files/figure-gfm/ca_country_count-1.png)
+![Figure 2](002_kb_rp_coverage_files/figure-gfm/ca_country_count-1.png)<!-- is the 'derived from wos_b_2019...' line really necessary in this file?
+only relevant for people with wos-access who know the tables-->
 
-After obtaining aggregated article counts at the level of countries and journals from the Web of Science, Crossref’s REST API was queried using the rcrossref package [@rcrossref] to retrieve publisher and journal titles.
+After obtaining aggregated article counts at the level of countries and journals from the Web of Science, Crossref’s REST API was queried using the rcrossref package [@rcrossref] to retrieve publisher <!--'publisher names' or 'publishers'--> and journal titles.
 To control for developments of the publishing market resulting in name changes of publishers or journal titles over time, only the most frequent field name was used from the API facet counts.
-Drawing on ISSN-L, a journal identifier that links between different media versions of a journal, ISSN variants were obtained to improve the retrieval.
-Next, the open access status per journal was identified and added to the dataset using the ISSN-Gold -OA 3.0 list [@issn].
+<!--most frequent name amongst all considered articles or in the whole of crossref?
+for this specific journal or always the same name over all journals?--> Drawing on ISSN-L, a journal identifier that links between different media versions of a journal, ISSN variants were obtained to improve the retrieval.
+Next, the open access status per journal was identified and added to the dataset using the ISSN-Gold-OA 3.0 list [@issn].
 Again, ISSN-L were used for the matching procedure.
@@
 - `journal_publisher_14_18`: Global journal and publisher data 2014-18
-- `rp_jn_14_18.csv`: Corresponding author country affiliations per publisher and journal 2014-18
+- `rp_jn_14_18`: Corresponding author country affiliations per publisher and journal 2014-18
 
-They are provided as comma-separated values (csv) files and Excel spreadsheets.
+They are provided as comma-separated values (csv) files and Excel spreadsheets (xlsx).
 
@@
 Table 1 describes the data variables in detail.
-We were able to retrieve publisher and journal titles from Crossref for `r publisher_df %>% filter(!is.na(publisher)) %>% distinct(issn_l) %>% nrow() / publisher_df %>% distinct(issn_l) %>% nrow() * 100` %
+We were able to retrieve publisher and journal title from Crossref for `r publisher_df %>% filter(!is.na(publisher)) %>% distinct(issn_l) %>% nrow() / publisher_df %>% distinct(issn_l) %>% nrow() * 100` %
  of all journals, representing `r publisher_df %>% filter(!is.na(publisher)) %>% .$articles %>% sum() / sum(publisher_df$articles) * 100` % of the overall article volume.
@@
 In total, the dataset provides information for `r sum(publisher_df$articles)` records indexed in the Web of Science with journal information in the five-years period 2014-2018.
-
-Data Schema `journal_publisher_14_18`:
+ <!-- are only journals with valid issn considered? Are there journals in the dataset without issn-l?-->
+<!--what about the 9% that could not be matched with crossref? are they excluded or included with WoS information /without journal information?-->
+Data Schema `journal_publisher_14_18`: <!--mention the abbreviation 'KB Web of Science' before somewhere when talking about KB-->
 
@@
 | ISSN-GOLD_OA List 3.0                                                                     |
-| `issn_l`           | Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.
+| `issn_l`           | Linking ISSN, a journal id that groups different media of the same serial publication, e.g. ISSN for print with electronic ISSN.
 | CIEPS                                                                                          |
@@
 | ISSN-GOLD_OA List 3.0                                                                                                                              |
-| `issn_l`           | Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.
+| `issn_l`           | Linking ISSN, a journal id that groups different media of the same serial publication, e.g. ISSN for print with electronic ISSN.
 | CIEPS                                                                                                                                                      |
@@
 The following table  shows the overall market share for the ten largest publishers.
-The following figure presents a breakdown by journal business model, differentiating between subscription journals and fully open access.
+The following figure presents a breakdown by journal business model, differentiating between subscription journals and fully open access journals.
 
@@
 
-![Figure 3](figure/global_publisher_share.png)
+![Figure 3](figure/global_publisher_share.png)<!-- Warum ist ab hier bei den Grafiken das Funding mit angegeben, bei den vorherigen jedoch nicht?-->
 
@@
 In total, ten publishers accounted for around two-thirds of articles published between 2014 - 2018.
-The Big 3, Elsevier, Springer Nature and Wiley, secured a market share of 46 %.
+The Big Three, Elsevier, Springer Nature and Wiley, secured a market share of 46 %.
 However, further analysis suggests large differences between subscription and fully open access journals: The market for fully open access journals seems to be less dominated by major publishing houses.
@@
   labs(
-    title = "Who did publish in scholarly journals between 2014 - 2018?",
+    title = "Who published in scholarly journals between 2014 - 2018?",
     x = "",
@@
     caption = "Data Sources:  Web of Science in-house database Competence Centre for Bibliometrics for German Science.
+    Publisher names were disambiguated using Crossref metadata.
     Funding: Federal Ministry of Education and Research of Germany (BMBF), Project OAUNI (01PU17023A)",
@@
 The flipped x-axis shows the top 10 countries of affiliations in terms of articles published, while the y-axis depicts the percentage per publisher.
-The figure indicates, which countries were of particularly importance to a publisher.
+The figure indicates, which countries were particularly important for a publisher.
 Furthermore, it suggests large variations across publishers.
 Take for instance the publisher Royal Society of Chemistry.
-Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(articles)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
+Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was significantly lower with `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(articles)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
 
@@
 However, there can be a considerable time-lag between online-first publication and issue inclusion.
+<!--mention also selection bias of WoS in terms of countries, ie.
+publications from EU, northern america much better covered than other continents-->
 
-Moreover, as any other bibliometric databases, the data structure can be subject of change.
-In particular, our pretests suggest that the Web of Science author role field used to identify corresponding authorships has undergone a change recently, starting to covering more than one reprint resp.
-corresponding author per article.
+Moreover, as is the case with any other bibliometric database, the data structure can be subject to change.
+In particular, our pretests suggest that the Web of Science author role field used to identify corresponding authorships has undergone a change recently, starting to cover more than one reprint or corresponding author per article in 2016.
 
-The ISSN GOLD-OA list [@issn] was used to determine fully open access journals.
+The ISSN GOLD-OA list <!--add version ?--> [@issn] was used to determine fully open access journals.
 However, it does not contain information when a subscription-based journal became fully open access.
-The figures should be therefore interpreted as numbers of articles available in current fully open access journals, instead of published in a fully open access journal at the time of publication.
+The figures should therefore be interpreted as numbers of articles available in current fully open access journals, instead of published in a fully open access journal at the time of publication.
 Open access article counts are therefore upper estimates.
+<!-- Is it really upper estimates?
+What about reverse flip?-->
 
 In the end, we present a novel bibliometric dataset representing the journal publishing market with a particular focus on open access.
-Although access to the underlying data infrastructure is restricted, we decided to openly share the aggregated dataset as part of a research compendium to demonstrate the technical reproducibility of this work.
+Although access to the underlying data infrastructure is restricted, we decided to openly share the aggregated dataset as part of a research compendium to demonstrate <!-- ensure instead of demonstrate?--> the technical reproducibility of this work.
 

### 2019-10-21 analysis/paper.Rmd

@@
 2. Coverage analysis of reprint authors in the WoS-KB for the period 2014 - 2018 ([`002_kb_rp_coverage.Rmd`](analysis/002_kb_rp_coverage.Rmd)).
-3. Steps to obtain and compile the dataset about corresponding author country affiliations 2014 - 2018([`003_kb_fetch_ca.Rmd`](analysis/003_kb_fetch_ca.Rmd)) 
+3. Steps to obtain and compile the dataset about corresponding author country affiliations 2014 - 2018 ([`003_kb_fetch_ca.Rmd`](analysis/003_kb_fetch_ca.Rmd))
 4. Steps to obtain and compile the dataset about the global publisher output 2014 - 2018 ([`004_kb_fetch_publisher.Rmd`](analysis/004_kb_fetch_publisher.Rmd)).
@@
 - `journal_publisher_14_18`: Global journal and publisher data 2014-18
-- `rp_jn_14_18`: Corresponding author country affiliations per publisher and journal 2014-18
+- `rp_publisher_14_18.csv`: Corresponding author country affiliations per publisher and journal 2014-18
 
@@
 ```{r}
-rp_df <- readr::read_csv("../data/rp_jn_14_18.csv")
+rp_df <- readr::read_csv("../data/rp_publisher_14_18.csv")
 ```
 
-The file named `rp_jn_14_18` provides a breakdown of journal information aggregated by countries of affiliation.
+The dataset named `rp_publisher_14_18` provides a breakdown of journal information aggregated by  countries of affiliation.
 Only corresponding authorships, indicated by the role "reprint author" in the Web of Science database, were taken into consideration.
@@
 
-Data
-Schema:
+Data Schema:
 
-| Variable           | Description                                                                                                                                                                      | Source                                                                                                                                                     |
-| :----------------- | :------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- | :--------------------------------------------------------------------------------------------------------------------------------------------------------- |
-| `issn_wos`         | ISSN, a standardized journal id.
-| KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                |
-| `country_code`     | Country of affiliation corresponding author, represented as ISO 3 code                                                                                                           | KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode`                                                                              |
-| `publication_year` | Year of publication, obtained from KB Web of Science                                                                                                                             | KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                              |
-| `articles`         | Number of original articles and reviews published.
-Whole counting where internationally co-located corresponding authorships were assigned to each contributing country equally.
-| KB Web of Science: Grouped counts over `wos_b_2019.issues.issn`, `wos_b_2019.d_items_authors_institutions.inst_countrycode` and `wos_b_2019.items.pubyear` |
-| `journal_title`    | Most frequently used journal title in terms of articles published between 2014 - 2018.
-If missing, the journal was not indexed in Crossref                                       | Crossref                                                                                                                                                   |
-| `publisher`        | Most frequently used publisher name in terms of articles published between 2014 - 2018.
-If missing, the journal was not indexed in Crossref                                      | Crossref                                                                                                                                                   |
-| `oa_journal`       | Is the journal publishing all articles open access without delay (full open access)?
-| ISSN-GOLD_OA List 3.0                                                                                                                              |
-| `issn_l`           | Linking ISSN, a journal id that groups different media of the same serial publication, e.g. ISSN for print with electronic ISSN.
-| CIEPS                                                                                                                                                      |
+|Variable           |Description                                                                                                                                 |Source                                                                        |
+|:------------------|:-------------------------------------------------------------------------------------------------------------------------------------------|:-----------------------------------------------------------------------------|
+|`country_code`     |Country of affiliation corresponding author, represented as ISO 3 code                                                                      |KB Web of Science: `wos_b_2019.d_items_authors_institutions.inst_countrycode` |
+|`publication_year` |Year of publication, obtained from KB Web of Science                                                                                        |KB Web of Science: `wos_b_2019.items.pubyear`                                 |
+|`publisher`        |Most frequently used publisher name in terms of articles published between 2014 - 2018.
+If missing, the journal was not indexed in Crossref |Crossref                                                                      |
+|`oa_journal`       |Is the journal publishing all articles open access without delay (full open access)?
+|ISSN GOLD OA List V3                                                          |
+|`n_publications`   |Number of original articles and reviews published                                                                                           |Aggregate data                                                                |
+|`n_journals`       |Number of distinct journals with corresponding publication output                                                                           |Aggregate data
 
@@
 Take for instance the publisher Royal Society of Chemistry.
-Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was significantly lower with `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(articles)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
+Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was significantly lower with `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(n_publications)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
 

### 2019-10-21 analysis/paper.Rmd
fix wording

@@
   labs(
-    title = "Who did publish in subscription-based journals from major publishers between 2014 - 2018?",
+    title = "Who published in subscription-based journals from major publishers between 2014 - 2018?",
     x = "Country of Affiliation (Top 10)",

### 2019-10-21 analysis/paper.Rmd
addressing @Ahobert comments and questions #1

@@
 In addition to university-specific data, the German Open Access Monitor provides breakdowns by publisher and journal and includes publications from non-university research institutions based in Germany [@oa_monitor].
+In its [Publisher Analytics Reports](https://clarivate.com/webofsciencegroup/solutions/publisher-analytics-reports/) Clarivate Analtics presents the performance of publishers across subjects, institutions and countries based on Web of Science data.
 
@@
 
-Following @Palzenberger_2015, data were retrieved from the in-house Web of Science database maintained by the German Competence Center for Bibliometrics.
+Following @Palzenberger_2015, data were retrieved from the in-house Web of Science database maintained by the German Competence Center for Bibliometrics (KB Web of Science).
 The following data were obtained and aggregated:
 
-- Web of Science collections SCI, SSCI and AHCI <!-- why only these collections?-->
+- Web of Science core collections SCI, SSCI and AHCI 
 - Article types Original Articles and Reviews
@@
 
-Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently: <!-- maybe stress that before 2016 only one unique corresponding author was allowed?--> We found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018 (see Figure 1).
+Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently: We found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018.
+However, Figure 1 suggests that indexing of more than one corresponding author started in 2016.
 Moreover, a 5% of corresponding authors author was internationally co-located (see Figure 2).
-We therefore<!--delete therefore?--> decided to use whole counting where every original article or review was counted once per country of affiliation of the corresponding author(s).
+We therefore decided to use whole counting where every original article or review was counted once per country of affiliation of the corresponding author(s).
 
@@
 
-After obtaining aggregated article counts at the level of countries and journals from the Web of Science, Crossref’s REST API was queried using the rcrossref package [@rcrossref] to retrieve publisher <!--'publisher names' or 'publishers'--> and journal titles.
-To control for developments of the publishing market resulting in name changes of publishers or journal titles over time, only the most frequent field name was used from the API facet counts.
-<!--most frequent name amongst all considered articles or in the whole of crossref?
-for this specific journal or always the same name over all journals?--> Drawing on ISSN-L, a journal identifier that links between different media versions of a journal, ISSN variants were obtained to improve the retrieval.
+After obtaining aggregated article counts at the level of countries and journals from the Web of Science, Crossref’s REST API was queried using the rcrossref package [@rcrossref] to retrieve publishers and journal titles.
+To control for developments of the publishing market resulting in name changes of publishers or journal titles over time, only the most frequent field name for the specific journal was used from the API facet counts.
+Drawing on ISSN-L, a journal identifier that links between different media versions of a journal, ISSN variants were obtained to improve the retrieval.
 Next, the open access status per journal was identified and added to the dataset using the ISSN-Gold-OA 3.0 list [@issn].
@@
  investigated journals with `ISSN-L` were identified as fully open access journals.
-In total, the dataset provides information for `r sum(publisher_df$articles)` records indexed in the Web of Science with journal information in the five-years period 2014-2018.
+In total, the dataset provides information for `r sum(publisher_df$articles)` records indexed in the Web of Science with journal information in the five-years period 2014-2018..
  <!-- are only journals with valid issn considered? Are there journals in the dataset without issn-l?-->
-<!--what about the 9% that could not be matched with crossref? are they excluded or included with WoS information /without journal information?-->
-Data Schema `journal_publisher_14_18`: <!--mention the abbreviation 'KB Web of Science' before somewhere when talking about KB-->
+<!--what about the 9% that could not be matched with crossref? are they excluded or included with WoS information /without journal information?-- per Wos ISSN-->
+Data Schema `journal_publisher_14_18`:
 
@@
 First, the Web of Science is selective in terms of journal inclusion and is not covering the whole journal portfolio of the investigated publishers.
+There is also evidence about a considerable indexing lag of non-English journal publications relevant to domestic research [@Lopez_Pineiro_2014].
 Although publisher names were disambiguated, journal title transfers between publishers were not tracked.
@@
 However, there can be a considerable time-lag between online-first publication and issue inclusion.
-<!--mention also selection bias of WoS in terms of countries, ie.
-publications from EU, northern america much better covered than other continents-->
 
@@
 
-The ISSN GOLD-OA list <!--add version ?--> [@issn] was used to determine fully open access journals.
+The ISSN GOLD-OA list [@issn] was used to determine fully open access journals.
 However, it does not contain information when a subscription-based journal became fully open access.
@@
 In the end, we present a novel bibliometric dataset representing the journal publishing market with a particular focus on open access.
-Although access to the underlying data infrastructure is restricted, we decided to openly share the aggregated dataset as part of a research compendium to demonstrate <!-- ensure instead of demonstrate?--> the technical reproducibility of this work.
+Although access to the underlying data infrastructure is restricted, we decided to openly share the aggregated dataset as part of a research compendium to ensure the technical reproducibility of this work.
 

### 2019-10-22 analysis/001_kb_rp_pretest.Rmd

@@
 Country information about corresponding authors play a crucial role in open access funding [@Schimmer_2015].
-The Web of Science contains data about this author role including affiliations, making it a valuable source to determine the productivity of corresponding authors per country ad journal.
+The Web of Science contains data about this author role including affiliations, making it a valuable source to determine the productivity of corresponding authors per country and journal.
 
@@
 
+<!--  why is issue of ex. 1 (multiple affiliations with differing author roles) not mentioned here?
+
+Maybe add something like
+
+4. Multiple country affiliations of the same author are listed with differing author roles
+
+The first three issues are of a more conceptual nature, all leading to a situation where one article is associated with multiple countries of affiliation.
+One has to decide upon a way of handling them which has to be considered at the time of interpretation of the results.
+Two options for the mentioned issues are whole counting or fractional counting.
+For whole counting, each article is counted fully for any country of affiliation of any listed corresponding author, so that the number of articles in the single categories do not necessarily add up to the total number of articles any more.
+Fractional counting assigns a fraction of one article count to each corresponding author, leading to possible non-integer total article counts.
+The second issue is not much of a problem in our setting, if both authors are situated in the same country, since we are interested in the country of affiliation associated with an article.
+If, however, the corresponding authors' countries of affiliation differ as in issue 3, we again have to decide upon a counting scheme: do we count each country only once per article or multiple times if multiple corresponding authors work there?
+The fourth issue is rather a problem of data consistency.
+-->
+
 ## Example 1
 
-A corresponding author lists at least two different countries, e.g., `UT=000400754000138`
+Potential issues: A corresponding author lists at least two different countries, e.g., `UT=000400754000138`.
+Additionally, only one affiliation is tagged as `RP` in `wos_b_2019.items_author_institutions`.
 
@@
 
-Potential issues: Only one affiliation is tagged as `RP` in `wos_b_2019.items_author_institutions`.
-
 ## Example 2
 
-More than one corresponding author is listed. For `UT_EID=000372645900002`, the Web of Science lists seven corresponding authors and four reprint adresses.
+Potential issues: More than one corresponding author is listed.
+For `UT_EID=000372645900002`, the Web of Science lists seven corresponding authors and four reprint adresses.
+Furthermore, as in Example 1, only one affiliation is tagged as `RP` in `wos_b_2019.items_author_institutions`.
 
@@
 
-Potential issue: As in Example 1, Only one affiliation is tagged as `RP` in `wos_b_2019.items_author_institutions`.
-
 ## Example 3
@@
 For `UT_EID=000372645900002`, the Web of Science lists two corresponding authors, one affilated with a Spanish, the other with a Portuguese institution.
+As in Examples 1 and 2, only one affiliation is tagged with RP in `wos_b_2019.items_author_institutions`.
+Moreover, checking the WoS raw data reveals that the RP-author `14734495` is not affiliated with an Portuguese institution at all.
+<!--when I run the query it only shows 10 records, none of them has fk_author 14734495-->
 
@@
 
-Potential issues: As in Example 1 and 2, only one affiliation is tagged with RP in `wos_b_2019.items_author_institutions`.
-Moreover, checking the WoS raw data reveals that the RP-author `14734495` is not affiliated with an Portuguese institution at all.
-
 ## SQL strategy to circumvent these issues
@@
 
-1. Get all `FK_AUTHORS` tagged as `RP`, and store the result `rp_author_table`
-2. From authors listed in `rp_author_table` obtain all countries of affilation and store them in `rp_countries_table`
+1. Get all `FK_AUTHORS` tagged as `RP`, and store the result as `rp_author_table`
+2. From authors listed in `rp_author_table` obtain all countries of affiliation and store them in `rp_countries_table`
 3. Aggregate `rp_countries_table`
 
-#### 1. Get all `FK_AUTHORS` tagged as `RP`, and store the result `rp_author_table`
+#### 1. Get all `FK_AUTHORS` tagged as `RP`, and store the result as `rp_author_table`
 

### 2019-10-22 analysis/001_kb_rp_pretest.Rmd
comment to data consistency #2
check author id and re-render notebook #2

@@
 <!--  why is issue of ex. 1 (multiple affiliations with differing author roles) not mentioned here?
+
+because I was not aware of this issue before. Let's wait and see how the table will be changed after the KB realised this data consistency issue.
 
@@
 As in Examples 1 and 2, only one affiliation is tagged with RP in `wos_b_2019.items_author_institutions`.
-Moreover, checking the WoS raw data reveals that the RP-author `14734495` is not affiliated with an Portuguese institution at all.
-<!--when I run the query it only shows 10 records, none of them has fk_author 14734495-->
+Moreover, checking the WoS raw data reveals that the RP-author `15145104` is not affiliated with an Portuguese institution at all.
 

### 2019-10-27 analysis/paper.Rmd
add colleen's review

@@
 However, funders and national library consortia often lack access to bibliometric data, which allow for determining the publication output per publisher, journal and journal business model across countries.
-Here, by obtaining country affiliations from the Web of Science at the level of journals, and combining them with normalized publisher and open access status information, we created a dataset about the productivity of corresponding authors for the period 2014-2018.
+Here, by obtaining country affiliations from the Web of Science, and combining them with normalized publisher and open access status information for journals, we created a dataset about the productivity of corresponding authors for the period 2014-2018.
 This dataset is expected to have essential value for the quantitative understanding of scholarly publishing.
@@
 Major bibliometric databases including the Web of Science, Scopus and Dimension have started to integrate open access status information.
-In these databases, Unpaywall data is often used to identify openly available full-texts at the article-level, while the Directory of Open Access Journals provides information about if a journal is fully open access.
+In these databases, Unpaywall data is often used to identify openly available full-texts at the article-level, while the Directory of Open Access Journals provides information on whether a journal is fully open access.
 In its most recent edition, the Leiden Ranking presents open access indicators at the level of universities [@leiden].
@@
 Here, we build an open dataset on the global productivity of corresponding authors.
-Drawing on Palzenberger's (2015) previous work, we not only created an updated dataset about the number and proportion of scholarly articles per country of affiliation from corresponding authors.
-The dataset also allows for a breakdown by journal, including information whether it was operated as a subscription or fully open access journal, and by publishers.
+Drawing on Palzenberger's (2015) previous work, we created an updated dataset that reveals the number and proportion of scholarly articles published per country, according to the corresponding author affiliation.
+The dataset also allows for a global breakdown by journal, including information whether it was operated as a subscription or fully open access journal, and by publishers.
 We constructed the dataset using the quality-assured Web of Science in-house database from the German Competence Center for Bibliometrics.
@@
 
-The data does not only provide background information for negotiations in the context of transformative agreements, but can also be used in the quantitative study of the scholarly publishing industry and the open access uptake across countries.
-For instance, national transformative agreements with large publishers will increase opportunities to publish open access.
-The existence of such agreements will therefore likely have an impact on the open access share of a country with such agreements.
+The data not only provides background information for negotiations in the context of transformative agreements, but can also be used in the quantitative study of the scholarly publishing industry and the uptake of open access publishing across countries.
+For instance, national transformative agreements with large publishers will increase authors'  opportunities to publish open access, and the existence of such agreements is therefore likely to have an impact on the open access share of a country with such agreements.
 To demonstrate the dataset's significance, we will present two new findings using the dataset: a global map of publishers' market shares, and an analysis of country of affiliation by publisher.
@@
 Like @Palzenberger_2015, the above-listed selection reflects data needs in the context of open access negotiations between library consortia and publishers.
-The restriction to the article types original article and review is in accordance with most open access contracts where other types of journal content like letters or meeting abstracts are not considered (see for instance @Sander_2019).
+The restriction to the article types original article and review is in accordance with most open access contracts where other types of journal content, such as letters or meeting abstracts, are not considered (see for instance @Sander_2019).
 
-To allocate open access costs to institutions, @Schimmer_2015 proposed to use affiliations from corresponding authors, presuming that corresponding author affiliations can be counted straight.
+To allocate open access costs to institutions, @Schimmer_2015 proposed using affiliations from corresponding authors, presuming that corresponding author affiliations can be counted straight.
 Under this counting method, only one institution or one country receives full credit for a publication [@Huang_2011].
-However, there have been some concerns with using corresponding authorship’s data from bibliometric databases for cost analysis recently.
+However, there have been some concerns raised over using corresponding authorship’s data from bibliometric databases for cost analysis recently.
 @Gumpenberger_2018 noted two potential issues: multiple corresponding author affiliations and multiple corresponding authors per article.
 
-Indeed, the Web of Science has begun to systematically keep track of more than one corresponding author recently: We found that the Web of Science did record more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018.
+Indeed, the Web of Science has recently begun to systematically keep track of more than one corresponding author per paper: We have found that the Web of Science recorded more than one corresponding author for 6.2% of indexed articles published between 2014 - 2018.
 However, Figure 1 suggests that indexing of more than one corresponding author started in 2016.
-Moreover, a 5% of corresponding authors author was internationally co-located (see Figure 2).
-We therefore decided to use whole counting where every original article or review was counted once per country of affiliation of the corresponding author(s).
+Moreover, 5% of corresponding authors had multiple affiliations from different countries (see Figure 2).
+We therefore decided to use whole counting in which every original article or review was counted once per country of affiliation of the corresponding author(s).
 
@@
 
-## Data and Code availability
+## Data and code availability
 
 The underlying source code in R and SQL for both testing the retrieval and compiling the datasets are openly shared as executable reports, making our data work more reproducible and transparent.
-The data analytics work, which was organized as research compendium [@Marwick_2017] using the holepunch R package [@holepunch], is accessible at <https://github.com/subugoe/oa2020cadata>.
+The data analytics work, which was organized as a research compendium [@Marwick_2017] using the holepunch R package [@holepunch], is accessible at <https://github.com/subugoe/oa2020cadata>.
 
@@
 
-1. Initial exploration about how corresponding authorships are represented in the WoS-KB ([`001_kb_rp_pretest.Rmd`](analysis/001_kb_rp_pretest.Rmd))
+1. Initial exploration on how corresponding authorships are represented in the WoS-KB ([`001_kb_rp_pretest.Rmd`](analysis/001_kb_rp_pretest.Rmd))
 2. Coverage analysis of reprint authors in the WoS-KB for the period 2014 - 2018 ([`002_kb_rp_coverage.Rmd`](analysis/002_kb_rp_coverage.Rmd)).
-3. Steps to obtain and compile the dataset about corresponding author country affiliations 2014 - 2018 ([`003_kb_fetch_ca.Rmd`](analysis/003_kb_fetch_ca.Rmd))
-4. Steps to obtain and compile the dataset about the global publisher output 2014 - 2018 ([`004_kb_fetch_publisher.Rmd`](analysis/004_kb_fetch_publisher.Rmd)).
+3. Steps to obtain and compile the dataset relative to corresponding author country affiliations 2014 - 2018 ([`003_kb_fetch_ca.Rmd`](analysis/003_kb_fetch_ca.Rmd))
+4. Steps to obtain and compile the dataset relative to the global publisher output 2014 - 2018 ([`004_kb_fetch_publisher.Rmd`](analysis/004_kb_fetch_publisher.Rmd)).
 
@@
 Table 1 describes the data variables in detail.
-We were able to retrieve publisher and journal title from Crossref for `r publisher_df %>% filter(!is.na(publisher)) %>% distinct(issn_l) %>% nrow() / publisher_df %>% distinct(issn_l) %>% nrow() * 100` %
+We were able to retrieve publisher and journal title from Crossref for `r publisher_df %>% filter(!is.na(publisher)) %>% distinct(issn_l) %>% nrow() / publisher_df %>% distinct(issn_l) %>% nrow() * 100`%
  of all journals, representing `r publisher_df %>% filter(!is.na(publisher)) %>% .$articles %>% sum() / sum(publisher_df$articles) * 100` % of the overall article volume.
@@
  investigated journals with `ISSN-L` were identified as fully open access journals.
-In total, the dataset provides information for `r sum(publisher_df$articles)` records indexed in the Web of Science with journal information in the five-years period 2014-2018..
+In total, the dataset provides information for `r sum(publisher_df$articles)` records indexed in the Web of Science with journal information in the five-year period 2014-2018.
+Journals that could not be matched with the `ISSN-L` or Crossref are marked with `NA`.
+
 <!-- are only journals with valid issn considered? Are there journals in the dataset without issn-l?-->
@@
 The following table  describes the variables in detail.
-Overall, 99 % of records representing original articles and reviews in the period 2014 - 2018 had affiliation information for corresponding authors at the country-level.
+Overall, 99% of records representing original articles and reviews in the period 2014 - 2018 had affiliation information for corresponding authors at the country-level.
 
@@
 
-Based on the dataset `journal_publisher_14_18` the global market share of publishing houses in terms of articles and reviews indexed in the Web of Science can be analyzed for the period 2014-18.
-The following table  shows the overall market share for the ten largest publishers.
+Based on the dataset `journal_publisher_14_18` the global market share of publishing houses as measured by articles and reviews indexed in the Web of Science can be analyzed for the period 2014-18.
+The following table shows the overall market share of the ten largest publishers.
 The following figure presents a breakdown by journal business model, differentiating between subscription journals and fully open access journals.
@@
   knitr::kable(
-    caption = "Publisher Market Share 2014 - 2018 in terms of articles and reviews indexed in the Web of Science",
+    caption = "Publisher Market Share 2014 - 2018 as measured by articles and reviews indexed in the Web of Science",
     col.names = c("Publisher", "Articles and Reviews published", "Percentage"),
@@
 The Big Three, Elsevier, Springer Nature and Wiley, secured a market share of 46 %.
-However, further analysis suggests large differences between subscription and fully open access journals: The market for fully open access journals seems to be less dominated by major publishing houses.
+However, further analysis suggests large differences between subscription and fully open access journals: the market of fully open access journals seems to be less dominated by major publishing houses.
 A notable exception is Springer Nature with its large fully open access portfolio including BioMed Central and the large mega-journals Scientific Reports and Nature Communications.
@@
 
-Using `rp_publisher_14_18.csv` the publisher market shares can be further broken down by country affiliation from corresponding authors.
+Using `rp_publisher_14_18.csv` the publisher market shares can be further broken down by country affiliation of corresponding authors.
 The following figure presents a global map of scholarly journal publishing represented in the Web of Science.
-Publications from institutions based in the European Union were summarized as "EU28".
-Together, all EU 28 member states constituted the largest market for journal publishing between 2014-18.
+Publications from institutions based in the European Union were summarized as "EU28", and, together, constitute the largest market for journal publishing between 2014-18.
+The figure further highlights the consistency of market dominance of the same set of large publishers across all regions of the world.
 
@@
 
-The following figure shows the country share per publisher in terms of original articles and reviews published in subscription journals using corresponding author affiliations.
-The flipped x-axis shows the top 10 countries of affiliations in terms of articles published, while the y-axis depicts the percentage per publisher.
-The figure indicates, which countries were particularly important for a publisher.
+The following figure shows the country share per publisher as measured by original articles and reviews published in subscription journals using corresponding author affiliations.
+The flipped x-axis shows the top 10 based on affiliations and articles published, while the y-axis depicts the percentage per publisher.
+The figure indicates, which countries were of particular relevance for a publisher.
 Furthermore, it suggests large variations across publishers.
-Take for instance the publisher Royal Society of Chemistry.
-Around every third article published in a subscription journal was submitted from an author based in China, while the overall share of China-based authors was significantly lower with `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(n_publications)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
+Take, for instance, the publisher Royal Society of Chemistry (RSC).
+Around every third article published in a subscription journal of RSC was submitted by an author based in China, while the overall share of China-based authors was significantly lower with `r rp_df %>% filter(oa_journal == FALSE) %>% group_by(country_code) %>% summarise(n = sum(n_publications)) %>% arrange(desc(n)) %>% mutate(prop = n /sum(n)) %>% filter(country_code == "CHN") %>% .$prop * 100` %.
 
@@
 
+### Discussion and outlook
+
+As country-level consortia increasingly embrace transformative agreements as a means to increase the proportion of their research published open access [^1], the dataset constitutes a highly useful resource for researchers and practitioners alike.
+When combined with information on a country’s own subscription and hybrid APC expenditure, bibliometric data about corresponding author affiliation benefit cost modelling for a fully open access future national consortia can derive key insights such as
+
+- What is the total research output of our authors?
+- How fast is our output growing?
+- What percentage of our authorship is as corresponding author (responsible for open access publishing costs)?
+- How much of our output is already published fully open access?
+- With which publishers do our authors most frequently publish?
+- How does the publisher distribution of our output relate to our subscription expenditure?
+- Would a shift to open access with publisher X cost more, less or the same as what we currently spend on subscriptions (and hybrid APCs)?
+- What value can we place on authorship and open access publishing services?
+
 ## Responsible use and limitations
@@
 Although the Web of Science database is a well-established bibliometric database it has certain limitations that need careful consideration when using and interpreting our data.
-First, the Web of Science is selective in terms of journal inclusion and is not covering the whole journal portfolio of the investigated publishers.
-There is also evidence about a considerable indexing lag of non-English journal publications relevant to domestic research [@Lopez_Pineiro_2014].
+First, the Web of Science is selective in terms of journal inclusion and does not covering the whole journal portfolio of the investigated publishers.
+There is also evidence about a considerable indexing gap of non-English journal publications relevant to domestic research [@Lopez_Pineiro_2014].
 Although publisher names were disambiguated, journal title transfers between publishers were not tracked.
@@
 Moreover, as is the case with any other bibliometric database, the data structure can be subject to change.
-In particular, our pretests suggest that the Web of Science author role field used to identify corresponding authorships has undergone a change recently, starting to cover more than one reprint or corresponding author per article in 2016.
+In particular, as mentioned above, our pretests suggest that the Web of Science author role field used to identify corresponding authorships has undergone a change recently, starting to cover more than one reprint or corresponding author per article in 2016.
 
@@
 However, it does not contain information when a subscription-based journal became fully open access.
-The figures should therefore be interpreted as numbers of articles available in current fully open access journals, instead of published in a fully open access journal at the time of publication.
-Open access article counts are therefore upper estimates.
-<!-- Is it really upper estimates?
+The figures should therefore be interpreted as numbers of articles available in current fully open access journals, rather than as published in a fully open access journal at the time of publication.
+<!-- Open access article counts are therefore upper estimates.
+Is it really upper estimates?
 What about reverse flip?-->
@@
 
+[^1]: See the ESAC registry for an up-to-date list of transformative agreements from national consortia <https://esac-initiative.org/about/transformative-agreements/agreement-registry/>
+
 ## References

### 2019-10-28 analysis/paper.Rmd
grammar fix

@@
 As country-level consortia increasingly embrace transformative agreements as a means to increase the proportion of their research published open access [^1], the dataset constitutes a highly useful resource for researchers and practitioners alike.
-When combined with information on a country’s own subscription and hybrid APC expenditure, bibliometric data about corresponding author affiliation benefit cost modelling for a fully open access future national consortia can derive key insights such as
+When combined with information on a country’s own subscription and hybrid APC expenditure, bibliometric data about corresponding author affiliation can help national consortia cost model for a fully open access future and derive key insights such as:
 

### 2019-10-29 analysis/paper.Rmd
typo

@@
 In addition to university-specific data, the German Open Access Monitor provides breakdowns by publisher and journal and includes publications from non-university research institutions based in Germany [@oa_monitor].
-In its [Publisher Analytics Reports](https://clarivate.com/webofsciencegroup/solutions/publisher-analytics-reports/) Clarivate Analtics presents the performance of publishers across subjects, institutions and countries based on Web of Science data.
+In its [Publisher Analytics Reports](https://clarivate.com/webofsciencegroup/solutions/publisher-analytics-reports/) Clarivate Analytics presents the performance of publishers across subjects, institutions and countries based on Web of Science data.
 
